---
title: "GSVA Recontextualization Evaluation"
author: "Andrew Chen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: 'united'
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(sigrecon)
PATH <- file.path(Sys.getenv("MLAB"), "projects/brcameta/projects/sig_recon")
```

# Loading Data
```{r}
# Load recon signatures
gsva_recon_sig_paths <- Sys.glob(file.path(PATH, "data/sigs/sciplex/recon/*gsva_*.rds"))
gsva_recon_sigs <- lapply(gsva_recon_sig_paths, readRDS)
names(gsva_recon_sigs) <- str_remove_all(basename(gsva_recon_sig_paths), "_gsva_recon")
names(gsva_recon_sigs) <- str_remove_all(names(gsva_recon_sigs), ".rds")

# Ground truth signatures
a549_sigs <- sigrecon::sciplex.a549
k562_sigs <- sigrecon::sciplex.k562
mcf7_sigs <- sigrecon::sciplex.mcf7

all_intersect_drugs <- purrr::reduce(lapply(names(gsva_recon_sigs), function(x) names(gsva_recon_sigs[[x]])),
                                 intersect)

a549_sigs <- a549_sigs[all_intersect_drugs]
k562_sigs <- k562_sigs[all_intersect_drugs]
mcf7_sigs <- mcf7_sigs[all_intersect_drugs]

# Filtering to shared pbs
gsva_recon_sigs <- lapply(gsva_recon_sigs, function(x) x[all_intersect_drugs])

a549_sigs_up <- lapply(a549_sigs, function(x) x$up)
k562_sigs_up <- lapply(k562_sigs, function(x) x$up)
mcf7_sigs_up <- lapply(mcf7_sigs, function(x) x$up)

a549_sigs_up_full <- lapply(a549_sigs, function(x) x$up_full)
k562_sigs_up_full <- lapply(k562_sigs, function(x) x$up_full)
mcf7_sigs_up_full <- lapply(mcf7_sigs, function(x) x$up_full)
```

```{r, message = FALSE, echo = FALSE}
# Define all cell line pairs and conditions
pairs <- c("a549_k562", "a549_mcf7", "k562_a549", "k562_mcf7", "mcf7_a549", "mcf7_k562")
conditions <- c("all", "control", "perturb")

# 1. Getting Data ==============================================================
metrics_data <- lapply(conditions, function(cond) {
  lapply(pairs, function(pair) {
    source_cell <- strsplit(pair, "_")[[1]][1]
    target_cell <- strsplit(pair, "_")[[1]][2]
    displaced_data <- count_displaced_genes(gsva_recon_sigs[[paste0(pair, "_", cond)]], get(paste0(target_cell, "_sigs_up")))
    ks_data <- v.ks.test(data_vecs = gsva_recon_sigs[[paste0(pair, "_", cond)]],
                         ref_vecs = get(paste0(target_cell, "_sigs_up_full")))
    list(
      jacc = v.jaccard(gsva_recon_sigs[[paste0(pair, "_", cond)]], get(paste0(target_cell, "_sigs_up"))),
      displaced = lapply(displaced_data, function(x) x$displaced) |> unlist() |> unname(),
      kept = lapply(displaced_data, function(x) x$not_displaced) |> unlist() |> unname(),
      k_d = ks_data$D_stats,
      k_p = ks_data$p_vals
    )
  }) |> setNames(paste0(pairs, "_", cond))
}) |> unlist(recursive = FALSE)

combined_recon_df <- dplyr::bind_rows(lapply(names(metrics_data), function(x) {
  source_name <- x
  metric_data <- metrics_data[[source_name]]
  data.frame(
    source = source_name,
    displaced = metric_data[["displaced"]],
    kept = metric_data[["kept"]],
    gene = names(gsva_recon_sigs[[source_name]]),
    jacc = metric_data[["jacc"]],
    k_d = metric_data[["k_d"]],
    k_p = metric_data[["k_p"]],
    recon = TRUE
  )
}))
```

```{r}
# 1. Getting Data ==============================================================
metrics_data <- lapply(pairs, function(pair) {
    source_cell <- strsplit(pair, "_")[[1]][1]
    target_cell <- strsplit(pair, "_")[[1]][2]
    displaced_data <- count_displaced_genes(get(paste0(source_cell, "_sigs_up")), get(paste0(target_cell, "_sigs_up")))
    ks_data <- v.ks.test(data_vecs = get(paste0(source_cell, "_sigs_up")),
                         ref_vecs = get(paste0(target_cell, "_sigs_up_full")))
    list(
      jacc = v.jaccard(get(paste0(source_cell, "_sigs_up")), get(paste0(target_cell, "_sigs_up"))),
      displaced = lapply(displaced_data, function(x) x$displaced) |> unlist() |> unname(),
      kept = lapply(displaced_data, function(x) x$not_displaced) |> unlist() |> unname(),
      k_d = ks_data$D_stats,
      k_p = ks_data$p_vals
    )
  }) |> setNames(pairs)

combined_no_recon_df <- dplyr::bind_rows(lapply(conditions, function(cond) {
  lapply(names(metrics_data), function(x) {
  source_name <- x
  metric_data <- metrics_data[[source_name]]
  data.frame(
    source = paste0(source_name, "_",cond),
    displaced = metric_data[["displaced"]],
    kept = metric_data[["kept"]],
    gene = names(gsva_recon_sigs[[paste0(source_name, "_",cond)]]),
    jacc = metric_data[["jacc"]],
    k_d = metric_data[["k_d"]],
    k_p = metric_data[["k_p"]],
    recon = FALSE
  )
  })
}))
```

```{r}
combined_aggregated_df <- merge(combined_no_recon_df %>% dplyr::select("source", "gene", "k_d", "jacc"), 
                                combined_recon_df %>% mutate(kept_alpha = kept/(kept+displaced)), 
                                by = c("source", "gene"), 
                                suffixes = c("_FALSE", "_TRUE")) %>% tibble
combined_aggregated_df
```

# Plots (All profiles) {.tabset}

## A549_K562 {.tabset}

K562 sigs recontextualized in A549 networks.

### KS
```{r}
combined_aggregated_df %>% 
  filter(source=="a549_k562_all") %>%
  ggplot(aes(x=k_d_FALSE, y=k_d_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("KS Distances in", "A549 Up"), 
     x = "KS Distance (Before)",
     y = "KS Distance (After)")
```

### Jaccard 
```{r}
combined_aggregated_df %>% 
  filter(source=="a549_k562_all") %>%
  ggplot(aes(x=jacc_FALSE, y=jacc_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("Jaccard Similarity in", "A549 Up"), 
     x = "Jaccard (Before)",
     y = "Jaccard (After)")
```

## A549_MCF7 {.tabset}

MCF7 sigs recontextualized in A549 networks.

### KS
```{r}
combined_aggregated_df %>% 
  filter(source=="a549_mcf7_all") %>%
  ggplot(aes(x=k_d_FALSE, y=k_d_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("KS Distances in", "A549 Up"), 
     x = "KS Distance (Before)",
     y = "KS Distance (After)")
```

### Jaccard 
```{r}
combined_aggregated_df %>% 
  filter(source=="a549_mcf7_all") %>%
  ggplot(aes(x=jacc_FALSE, y=jacc_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("Jaccard Similarity in", "A549 Up"), 
     x = "Jaccard (Before)",
     y = "Jaccard (After)")
```

## K562_A549 {.tabset}

A549 sigs recontextualized in K562 networks.

### KS
```{r}
combined_aggregated_df %>% 
  filter(source=="k562_a549_all") %>%
  ggplot(aes(x=k_d_FALSE, y=k_d_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("KS Distances in", "K562 Up"), 
     x = "KS Distance (Before)",
     y = "KS Distance (After)")
```

### Jaccard 
```{r}
combined_aggregated_df %>% 
  filter(source=="k562_a549_all") %>%
  ggplot(aes(x=jacc_FALSE, y=jacc_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("Jaccard Similarity in", "K562 Up"), 
     x = "Jaccard (Before)",
     y = "Jaccard (After)")
```

## K562_MCF7 {.tabset}

MCF7 sigs recontextualized in K562 networks.

### KS
```{r}
combined_aggregated_df %>% 
  filter(source=="k562_mcf7_all") %>%
  ggplot(aes(x=k_d_FALSE, y=k_d_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("KS Distances in", "K562 Up"), 
     x = "KS Distance (Before)",
     y = "KS Distance (After)")
```

### Jaccard 
```{r}
combined_aggregated_df %>% 
  filter(source=="k562_mcf7_all") %>%
  ggplot(aes(x=jacc_FALSE, y=jacc_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("Jaccard Similarity in", "K562 Up"), 
     x = "Jaccard (Before)",
     y = "Jaccard (After)")
```

## MCF7_A549 {.tabset}

A549 sigs recontextualized in MCF7 networks.

### KS
```{r}
combined_aggregated_df %>% 
  filter(source=="mcf7_a549_all") %>%
  ggplot(aes(x=k_d_FALSE, y=k_d_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("KS Distances in", "MCF7 Up"), 
     x = "KS Distance (Before)",
     y = "KS Distance (After)")
```

### Jaccard 
```{r}
combined_aggregated_df %>% 
  filter(source=="mcf7_a549_all") %>%
  ggplot(aes(x=jacc_FALSE, y=jacc_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("Jaccard Similarity in", "MCF7 Up"), 
     x = "Jaccard (Before)",
     y = "Jaccard (After)")
```

## MCF7_K562 {.tabset}

K562 sigs recontextualized in MCF7 networks.

### KS
```{r}
combined_aggregated_df %>% 
  filter(source=="mcf7_k562_all") %>%
  ggplot(aes(x=k_d_FALSE, y=k_d_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("KS Distances in", "MCF7 Up"), 
     x = "KS Distance (Before)",
     y = "KS Distance (After)")
```

### Jaccard 
```{r}
combined_aggregated_df %>% 
  filter(source=="mcf7_k562_all") %>%
  ggplot(aes(x=jacc_FALSE, y=jacc_TRUE)) +
  geom_point(aes(alpha=kept_alpha)) +
  geom_abline(color="black") +
  scale_alpha_continuous("Proportion of Source Kept") +
  labs(title = paste("Jaccard Similarity in", "MCF7 Up"), 
     x = "Jaccard (Before)",
     y = "Jaccard (After)")
```

# Unpaired
```{r}
combined_aggregated_df %>% 
  dplyr::filter(str_detect(source, "control")) %>%
  dplyr::group_by(source) %>%
  dplyr::summarize(kd_mean = mean(k_d_TRUE),
                   jacc_mean = mean(jacc_TRUE)) %>%
  dplyr::summarize(kd_mean = mean(kd_mean),
                   jacc_mean = mean(jacc_mean))
```


# Pairwise summary
```{r}
combined_aggregated_df$k_d_TRUE %>% mean
combined_aggregated_df$jacc_TRUE %>% mean
```

```{r}
test <- combined_aggregated_df %>% 
  mutate(Recon_Succ_Ks = k_d_TRUE > k_d_FALSE + 1e-2,
         Recon_Fail_Ks = k_d_TRUE < k_d_FALSE - 1e-2,
         Recon_Succ_Jacc = jacc_TRUE > jacc_FALSE + 1e-2,
         Recon_Fail_Jacc = jacc_TRUE < jacc_FALSE - 1e-2) %>% 
  group_by(source) %>% 
  summarize(Recon_Works_Jacc = sum(Recon_Succ_Jacc)/(sum(Recon_Succ_Jacc) + sum(Recon_Fail_Jacc)),
            Recon_Works_KS = sum(Recon_Succ_Ks)/(sum(Recon_Succ_Ks) + sum(Recon_Fail_Ks)),
            Jacc_Wilcox_pval = wilcox.test(jacc_TRUE, jacc_FALSE, paired = TRUE, alternative="greater")$p.value,
            Jacc_t_pval = t.test(jacc_TRUE, jacc_FALSE, paired=TRUE, alternative="greater")$p.value,
            KS_Wilcox_pval = wilcox.test(k_d_TRUE, k_d_FALSE, paired = TRUE, alternative="greater")$p.value,
            KS_t_pval = t.test(k_d_TRUE, k_d_FALSE, paired=TRUE, alternative="greater")$p.value)
test
```

```{r}
test %>% dplyr::mutate(profile = case_when(str_detect(source, "all") ~ "all",
                                           str_detect(source, "control") ~ "control",
                                           str_detect(source, "perturb") ~ "perturb")) %>% 
    dplyr::group_by(profile) %>%
    dplyr::summarize(prop_jacc = mean(Recon_Works_Jacc),
                     prop_ks = mean(Recon_Works_KS))
```

```{r}
test %>% dplyr::filter(str_detect(source, "control")) %>% dplyr::summarize(KS_meta_p = fishers_meta_p(KS_Wilcox_pval), Jacc_meta_p = fishers_meta_p(Jacc_Wilcox_pval))
```

