---
title: "Figures"
author: "Andrew Chen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: 'united'
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(sigrecon)
library(ggplot2)
library(ggbeeswarm)
library(arrow)
library(ComplexHeatmap)
library(circlize)
library(grid)

PROJECT_PATH <- file.path(Sys.getenv("MLAB"), "projects/brcameta/projects/sig_recon")
RESULTS_PATH <- file.path(PROJECT_PATH, "results/eval")

# Set to TRUE to recompute and overwrite saved objects; FALSE to load from saved files
do_save            <- FALSE   # summary tables (Unpaired / Paired results sections)
do_save_tahoe      <- FALSE   # Tahoe visualization plot objects
do_save_drugmatrix <- FALSE   # DrugMatrix visualization plot objects
do_save_perturbseq <- FALSE   # Perturb-seq visualization plot objects
do_save_sciplex    <- FALSE   # SciPlex visualization plot objects
```

```{r}
summarize_method <- function(df, method_name){

  per_source <- df %>%
    group_by(source) %>%
    summarise(
      jacc_p = wilcox.test(jacc_TRUE, jacc_FALSE,
                           paired = TRUE,
                           alternative = "greater")$p.value,

      nes_p = wilcox.test(NES_TRUE, NES_FALSE,
                          paired = TRUE,
                          alternative = "greater")$p.value,

      jacc_prop = mean(jacc_TRUE > jacc_FALSE),
      nes_prop = mean(NES_TRUE > NES_FALSE),
      .groups="drop"
    )

  tibble(
    method = method_name,
    metric = c("Jaccard","NES"),
    fisher_p = c(
      fishers_meta_p(per_source$jacc_p),
      fishers_meta_p(per_source$nes_p)
    ),
    prop = c(
      mean(per_source$jacc_prop),
      mean(per_source$nes_prop)
    )
  )
}
```

# Unpaired results {.tabset}

```{r unpaired-results, message=FALSE, warning=FALSE}
# ── Helper: load an RDS or return NULL ────────────────────────────────────────
load_rds <- function(dir, file) {
  path <- file.path(RESULTS_PATH, dir, file)
  if (file.exists(path)) readRDS(path) else NULL
}

# ── Normalise column names: no_change tables use NES/jacc instead of NES_TRUE/jacc_TRUE
normalize_eval_cols <- function(df) {
  if (!"NES_TRUE"  %in% names(df) && "NES"  %in% names(df)) df <- dplyr::rename(df, NES_TRUE  = NES)
  if (!"jacc_TRUE" %in% names(df) && "jacc" %in% names(df)) df <- dplyr::rename(df, jacc_TRUE = jacc)
  df
}

# ── Summarise one eval table → single-row tibble (NES_mean, jacc_mean) ───────
summarize_unpaired <- function(eval_table) {
  if (is.null(eval_table)) return(tibble(NES_mean = NA_real_, jacc_mean = NA_real_))
  eval_table <- normalize_eval_cols(eval_table)
  eval_table %>%
    dplyr::group_by(source) %>%
    dplyr::summarise(NES_mean  = median(NES_TRUE,  na.rm = TRUE),
                     jacc_mean = median(jacc_TRUE, na.rm = TRUE),
                     .groups = "drop") %>%
    dplyr::summarise(NES_mean  = median(NES_mean),
                     jacc_mean = median(jacc_mean))
}

# ── File map: list(dataset = list(regime = list(method = filename))) ──────────
file_map <- list(
  drugmatrix = list(
    ctrl = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_mean_eval_table.rds",
      pc_eigen    = "ctrl_projectcor_eigen_eval_table.rds",
      pc_gsva     = "ctrl_projectcor_gsva_eval_table.rds",
      networkprop = "ctrl_networkprop_eval_table.rds",
      orthos      = "ctrl_orthos_residual_eval_table.rds",
      scgpt       = NULL,
      stack       = NULL
    ),
    `10th` = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_10th_mean_eval_table.rds",
      pc_eigen    = "10th_projectcor_eigen_eval_table.rds",
      pc_gsva     = "10th_projectcor_gsva_eval_table.rds",
      networkprop = "ctrl_10th_networkprop_eval_table.rds",
      scgpt       = NULL,
      stack       = NULL
    ),
    `90th` = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_90th_mean_eval_table.rds",
      pc_eigen    = "90th_projectcor_eigen_eval_table.rds",
      pc_gsva     = "90th_projectcor_gsva_eval_table.rds",
      networkprop = "ctrl_90th_networkprop_eval_table.rds",
      scgpt       = NULL,
      stack       = NULL
    )
  ),
  `perturb-seq` = list(
    ctrl = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_mean_eval_table.rds",
      pc_eigen    = "ctrl_projectcor_eigen_eval_table.rds",
      pc_gsva     = "ctrl_projectcor_gsva_eval_table.rds",
      networkprop = "ctrl_networkprop_eval_table.rds",
      orthos      = NULL,
      scgpt       = NULL,
      stack       = "ctrl_stack_eval_table.rds"
    ),
    `10th` = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_10th_mean_eval_table.rds",
      pc_eigen    = "10th_projectcor_eigen_eval_table.rds",
      pc_gsva     = "10th_projectcor_gsva_eval_table.rds",
      networkprop = "ctrl_10th_networkprop_eval_table.rds",
      scgpt       = "scgpt_10th_eval_table.rds",
      stack       = "10th_stack_eval_table.rds"
    ),
    `90th` = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_90th_mean_eval_table.rds",
      pc_eigen    = "90th_projectcor_eigen_eval_table.rds",
      pc_gsva     = "90th_projectcor_gsva_eval_table.rds",
      networkprop = "ctrl_90th_networkprop_eval_table.rds",
      scgpt       = "scgpt_90th_eval_table.rds",
      stack       = "90th_stack_eval_table.rds"
    )
  ),
  sciplex = list(
    ctrl = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_mean_eval_table.rds",
      pc_eigen    = "ctrl_projectcor_eigen_eval_table.rds",
      pc_gsva     = "ctrl_projectcor_gsva_eval_table.rds",
      networkprop = "ctrl_networkprop_eval_table.rds",
      orthos      = "ctrl_orthos_residual_eval_table.rds",
      scgpt       = NULL,
      stack       = "stack_ctrl_eval_table.rds"
    ),
    `10th` = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_10th_mean_eval_table.rds",
      pc_eigen    = "10th_projectcor_eigen_eval_table.rds",
      pc_gsva     = "10th_projectcor_gsva_eval_table.rds",
      networkprop = "10th_networkprop_eval_table.rds",
      scgpt       = NULL,
      stack       = "stack_10th_eval_table.rds"
    ),
    `90th` = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_90th_mean_eval_table.rds",
      pc_eigen    = "90th_projectcor_eigen_eval_table.rds",
      pc_gsva     = "90th_projectcor_gsva_eval_table.rds",
      networkprop = "90th_networkprop_eval_table.rds",
      scgpt       = NULL,
      stack       = "stack_90th_eval_table.rds"
    )
  ),
  tahoe = list(
    ctrl = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_mean_eval_table.rds",
      pc_eigen    = "ctrl_projectcor_eigen_eval_table.rds",
      pc_gsva     = "ctrl_projectcor_gsva_eval_table.rds",
      networkprop = "ctrl_networkprop_eval_table.rds",
      orthos      = "ctrl_orthos_residual_eval_table.rds",
      scgpt       = NULL,
      stack       = "stack_ctrl_eval_table.rds"
    ),
    `10th` = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_10th_mean_eval_table.rds",
      pc_eigen    = "10th_projectcor_eigen_eval_table.rds",
      pc_gsva     = "10th_projectcor_gsva_eval_table.rds",
      networkprop = "network_prop_10th_eval_table.rds",
      scgpt       = NULL,
      stack       = "stack_10th_eval_table.rds"
    ),
    `90th` = list(
      no_change   = "no_change_eval_table.rds",
      mean        = "ctrl_90th_mean_eval_table.rds",
      pc_eigen    = "90th_projectcor_eigen_eval_table.rds",
      pc_gsva     = "90th_projectcor_gsva_eval_table.rds",
      networkprop = "network_prop_90th_eval_table.rds",
      scgpt       = NULL,
      stack       = "stack_90th_eval_table.rds"
    )
  )
)

# ── Method display labels (in desired row order) ──────────────────────────────
method_labels <- c(
  no_change   = "no-change",
  mean        = "mean",
  pc_eigen    = "project cor (eigen)",
  pc_gsva     = "project cor (gsva)",
  networkprop = "network prop",
  orthos      = "orthos",
  scgpt       = "scGPT",
  stack       = "stack"
)

# Methods that only exist for the ctrl regime (omitted from 10th/90th tables)
ctrl_only_methods <- c("orthos")

datasets        <- c("drugmatrix", "perturb-seq", "sciplex", "tahoe")
dataset_labels  <- c("drugmatrix", "perturbseq",  "sciplex", "tahoe")
methods         <- names(method_labels)
regimes         <- c("ctrl", "10th", "90th")

# ── Build one unpaired summary table for a given regime ──────────────────────
build_unpaired_table <- function(regime) {
  regime_methods <- if (regime == "ctrl") methods else methods[!methods %in% ctrl_only_methods]
  rows <- lapply(regime_methods, function(m) {
    vals <- lapply(seq_along(datasets), function(i) {
      ds   <- datasets[i]
      fname <- file_map[[ds]][[regime]][[m]]
      tbl   <- if (!is.null(fname)) load_rds(ds, fname) else NULL
      summarize_unpaired(tbl)
    })
    row <- tibble(method = method_labels[m])
    for (i in seq_along(datasets)) {
      dl <- dataset_labels[i]
      row[[paste0(dl, "_NES")]]  <- vals[[i]]$NES_mean
      row[[paste0(dl, "_Jacc")]] <- vals[[i]]$jacc_mean
    }
    row
  })
  bind_rows(rows) %>%
    dplyr::select(method,
                  drugmatrix_NES, drugmatrix_Jacc,
                  perturbseq_NES, perturbseq_Jacc,
                  sciplex_NES,    sciplex_Jacc,
                  tahoe_NES,      tahoe_Jacc)
}

csv_dir <- file.path(RESULTS_PATH, "summary_tables")

if (do_save) {
  unpaired_ctrl <- build_unpaired_table("ctrl")
  unpaired_10th <- build_unpaired_table("10th")
  unpaired_90th <- build_unpaired_table("90th")
  dir.create(csv_dir, showWarnings = FALSE, recursive = TRUE)
  write.csv(unpaired_ctrl, file.path(csv_dir, "unpaired_ctrl.csv"),  row.names = FALSE)
  write.csv(unpaired_10th, file.path(csv_dir, "unpaired_10th.csv"), row.names = FALSE)
  write.csv(unpaired_90th, file.path(csv_dir, "unpaired_90th.csv"), row.names = FALSE)
} else {
  unpaired_ctrl <- read.csv(file.path(csv_dir, "unpaired_ctrl.csv"))
  unpaired_10th <- read.csv(file.path(csv_dir, "unpaired_10th.csv"))
  unpaired_90th <- read.csv(file.path(csv_dir, "unpaired_90th.csv"))
}
```

## ctrl
```{r}
unpaired_ctrl
```

## 10th
```{r}
unpaired_10th
```

## 90th
```{r}
unpaired_90th
```

# Paired results {.tabset}

```{r paired-results, message=FALSE, warning=FALSE}
# ── Summarise one eval table → single-row tibble (KS_meta_p, Jacc_meta_p) ────
summarize_paired <- function(eval_table) {
  if (is.null(eval_table)) return(tibble(KS_meta_p = NA_real_, Jacc_meta_p = NA_real_))
  eval_table <- normalize_eval_cols(eval_table)
  eval_table %>%
    dplyr::group_by(source) %>%
    dplyr::summarize(
      Jacc_Wilcox_pval = wilcox.test(jacc_TRUE, jacc_FALSE,
                                     paired = TRUE, alternative = "greater")$p.value,
      KS_Wilcox_pval   = wilcox.test(NES_TRUE,  NES_FALSE,
                                     paired = TRUE, alternative = "greater")$p.value,
      .groups = "drop"
    ) %>%
    dplyr::summarize(
      KS_meta_p   = fishers_meta_p(pmax(KS_Wilcox_pval,   .Machine$double.eps)),
      Jacc_meta_p = fishers_meta_p(pmax(Jacc_Wilcox_pval, .Machine$double.eps))
    )
}

# Methods for paired results (no no-change)
paired_methods <- methods[methods != "no_change"]

# ── Build one paired summary table for a given regime ────────────────────────
build_paired_table <- function(regime) {
  regime_methods <- if (regime == "ctrl") paired_methods else paired_methods[!paired_methods %in% ctrl_only_methods]
  rows <- lapply(regime_methods, function(m) {
    vals <- lapply(seq_along(datasets), function(i) {
      ds    <- datasets[i]
      fname <- file_map[[ds]][[regime]][[m]]
      tbl   <- if (!is.null(fname)) load_rds(ds, fname) else NULL
      summarize_paired(tbl)
    })
    row <- tibble(method = method_labels[m])
    for (i in seq_along(datasets)) {
      dl <- dataset_labels[i]
      row[[paste0(dl, "_NES")]]  <- vals[[i]]$KS_meta_p
      row[[paste0(dl, "_Jacc")]] <- vals[[i]]$Jacc_meta_p
    }
    row
  })
  bind_rows(rows) %>%
    dplyr::select(method,
                  drugmatrix_NES, drugmatrix_Jacc,
                  perturbseq_NES, perturbseq_Jacc,
                  sciplex_NES,    sciplex_Jacc,
                  tahoe_NES,      tahoe_Jacc)
}

if (do_save) {
  paired_ctrl <- build_paired_table("ctrl")
  paired_10th <- build_paired_table("10th")
  paired_90th <- build_paired_table("90th")
  write.csv(paired_ctrl, file.path(csv_dir, "paired_ctrl.csv"),  row.names = FALSE)
  write.csv(paired_10th, file.path(csv_dir, "paired_10th.csv"), row.names = FALSE)
  write.csv(paired_90th, file.path(csv_dir, "paired_90th.csv"), row.names = FALSE)
} else {
  paired_ctrl <- read.csv(file.path(csv_dir, "paired_ctrl.csv"))
  paired_10th <- read.csv(file.path(csv_dir, "paired_10th.csv"))
  paired_90th <- read.csv(file.path(csv_dir, "paired_90th.csv"))
}

# ── Summarise one eval table → single-row tibble (KS_stat, Jacc_stat) ────────
summarize_paired_stat <- function(eval_table) {
  if (is.null(eval_table)) return(tibble(KS_stat = NA_real_, Jacc_stat = NA_real_))
  eval_table <- normalize_eval_cols(eval_table)
  if (!all(c("jacc_FALSE", "NES_FALSE") %in% names(eval_table))) {
    return(tibble(KS_stat = NA_real_, Jacc_stat = NA_real_))
  }
  eval_table %>%
    dplyr::group_by(source) %>%
    dplyr::summarize(
      Jacc_stat = wilcox.test(jacc_TRUE, jacc_FALSE,
                              paired = TRUE, alternative = "greater")$statistic,
      KS_stat   = wilcox.test(NES_TRUE,  NES_FALSE,
                              paired = TRUE, alternative = "greater")$statistic,
      .groups = "drop"
    ) %>%
    dplyr::summarize(
      KS_stat   = median(KS_stat),
      Jacc_stat = median(Jacc_stat)
    )
}

# ── Build one paired statistic table for a given regime ──────────────────────
build_paired_stat_table <- function(regime) {
  regime_methods <- if (regime == "ctrl") paired_methods else paired_methods[!paired_methods %in% ctrl_only_methods]
  rows <- lapply(regime_methods, function(m) {
    vals <- lapply(seq_along(datasets), function(i) {
      ds    <- datasets[i]
      fname <- file_map[[ds]][[regime]][[m]]
      tbl   <- if (!is.null(fname)) load_rds(ds, fname) else NULL
      summarize_paired_stat(tbl)
    })
    row <- tibble(method = method_labels[m])
    for (i in seq_along(datasets)) {
      dl <- dataset_labels[i]
      row[[paste0(dl, "_NES")]]  <- vals[[i]]$KS_stat
      row[[paste0(dl, "_Jacc")]] <- vals[[i]]$Jacc_stat
    }
    row
  })
  bind_rows(rows) %>%
    dplyr::select(method,
                  drugmatrix_NES, drugmatrix_Jacc,
                  perturbseq_NES, perturbseq_Jacc,
                  sciplex_NES,    sciplex_Jacc,
                  tahoe_NES,      tahoe_Jacc)
}

if (do_save) {
  paired_stat_ctrl <- build_paired_stat_table("ctrl")
  paired_stat_10th <- build_paired_stat_table("10th")
  paired_stat_90th <- build_paired_stat_table("90th")
  write.csv(paired_stat_ctrl, file.path(csv_dir, "paired_stat_ctrl.csv"),  row.names = FALSE)
  write.csv(paired_stat_10th, file.path(csv_dir, "paired_stat_10th.csv"), row.names = FALSE)
  write.csv(paired_stat_90th, file.path(csv_dir, "paired_stat_90th.csv"), row.names = FALSE)
} else {
  paired_stat_ctrl <- read.csv(file.path(csv_dir, "paired_stat_ctrl.csv"))
  paired_stat_10th <- read.csv(file.path(csv_dir, "paired_stat_10th.csv"))
  paired_stat_90th <- read.csv(file.path(csv_dir, "paired_stat_90th.csv"))
}
```

## ctrl {.tabset}

### p-value
```{r}
paired_ctrl
```

### test statistic
```{r}
paired_stat_ctrl
```

## 10th {.tabset}

### p-value
```{r}
paired_10th
```

### test statistic
```{r}
paired_stat_10th
```

## 90th {.tabset}

### p-value
```{r}
paired_90th
```

### test statistic
```{r}
paired_stat_90th
```

# Tahoe {.tabset}

## Control {.tabset}
```{r}
if (do_save_tahoe) {
  tahoe_no_change_df   <- readRDS(file.path(RESULTS_PATH, "tahoe/no_change_eval_table.rds"))
  tahoe_ctrl_mean_df   <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_mean_eval_table.rds"))
  tahoe_ctrl_net_df    <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_networkprop_eval_table.rds"))
  tahoe_ctrl_proj_e_df <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_projectcor_eigen_eval_table.rds"))
  tahoe_ctrl_proj_g_df <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_projectcor_gsva_eval_table.rds"))
  tahoe_orthos_df      <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_orthos_residual_eval_table.rds"))
  tahoe_stack_df       <- readRDS(file.path(RESULTS_PATH, "tahoe/stack_ctrl_eval_table.rds"))
}
```

```{r}
if (do_save_tahoe) {
  plot_df <- bind_rows(
    tahoe_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),

    tahoe_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_orthos_df %>%
      transmute(drug = gene, method = "Orthos", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_stack_df %>%
      transmute(drug = gene, method = "Stack", jacc = jacc_TRUE, NES = NES_TRUE),
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "tahoe/ctrl_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(tahoe_ctrl_mean_df, "Mean"),
    summarize_method(tahoe_ctrl_net_df, "NetworkProp"),
    summarize_method(tahoe_ctrl_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(tahoe_ctrl_proj_g_df, "ProjectCor (GSVA)"),
    summarize_method(tahoe_orthos_df, "Orthos"),
    summarize_method(tahoe_stack_df, "Stack")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "tahoe/ctrl_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +
  geom_violin(
    width = 0.6,
    alpha = 0.3,
    color = NA,
    trim = TRUE
  ) +
  ggbeeswarm::geom_quasirandom(
    width = 0.15,
    alpha = 0.15,
    size = 0.05,
    groupOnX = TRUE
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    shape = 21,
    size = 2.5,
    fill = "white"
  ) +
  geom_hline(
    data = baseline_means,
    aes(yintercept = yintercept),
    inherit.aes = FALSE,
    linetype = "dotted",
    color = "grey50",
    linewidth = 0.8
  ) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(
    limits = c("No change", "Mean", "NetworkProp", "ProjectCor (Eigen)",
               "ProjectCor (GSVA)", "Orthos", "Stack")
  ) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/tahoe_ctrl_unpaired.png"),
  width = 8,
  height = 5,
  dpi = 400
)
```

### Paired {.tabset}

#### Swarm
```{r}
if (do_save_tahoe) {
  swarm_df <- bind_rows(
    tahoe_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_orthos_df %>%
      transmute(drug = gene, method = "Orthos",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_stack_df %>%
      transmute(drug = gene, method = "Stack",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  swarm_long <- swarm_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(swarm_long, file.path(RESULTS_PATH, "tahoe/ctrl_paired_swarm_plot_df.rds"))
} else {
  swarm_long <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(swarm_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(width = 0.15, alpha = 0.15, size = 0.05, groupOnX = TRUE) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Orthos", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/tahoe_ctrl_paired_swarm.png"),
  width = 8, height = 5, dpi = 400)
```

## 10th Results {.tabset}
```{r}
if (do_save_tahoe) {
  tahoe_no_change_df   <- readRDS(file.path(RESULTS_PATH, "tahoe/no_change_eval_table.rds"))
  tahoe_ctrl_mean_df   <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_10th_mean_eval_table.rds"))
  tahoe_ctrl_net_df    <- readRDS(file.path(RESULTS_PATH, "tahoe/network_prop_10th_eval_table.rds"))
  tahoe_ctrl_proj_e_df <- readRDS(file.path(RESULTS_PATH, "tahoe/10th_projectcor_eigen_eval_table.rds"))
  tahoe_ctrl_proj_g_df <- readRDS(file.path(RESULTS_PATH, "tahoe/10th_projectcor_gsva_eval_table.rds"))
}
```

```{r}
if (do_save_tahoe) {
  plot_df <- bind_rows(
    tahoe_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),

    tahoe_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    dplyr::group_by(drug, method) %>%
    dplyr::summarise_all(mean) %>%
    dplyr::ungroup() %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "tahoe/ctrl_10th_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(tahoe_ctrl_mean_df, "Mean"),
    summarize_method(tahoe_ctrl_net_df, "NetworkProp"),
    summarize_method(tahoe_ctrl_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(tahoe_ctrl_proj_g_df, "ProjectCor (GSVA)")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "tahoe/ctrl_10th_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_10th_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_10th_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +

  geom_violin(
    width = 0.9,
    alpha = 0.35,
    color = NA
  ) +

  ggbeeswarm::geom_beeswarm(
    size = 0.1,
    alpha = 0.7,
    cex = 0.7
  ) +

  stat_summary(
    fun = median,
    geom = "point",
    shape = 21,
    size = 3,
    fill = "white"
  ) +

  geom_hline(
    data = baseline_means,
    aes(yintercept = yintercept),
    linetype = "dotted",
    color = "grey50",
    linewidth = 0.8
  ) +

  facet_wrap(
    ~metric,
    scales = "free_y"
  ) +

  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)")) + 
  labs(
    x = "",
    y = "Score"
  ) +

  theme_classic(base_size = 15) +

  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/tahoe_10th_unpaired.png"),
  width = 8,
  height = 5,
  dpi = 400
)
```

### Paired {.tabset}

#### Swarm
```{r}
if (do_save_tahoe) {
  swarm_df <- bind_rows(
    tahoe_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  swarm_long <- swarm_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(swarm_long, file.path(RESULTS_PATH, "tahoe/ctrl_10th_paired_swarm_plot_df.rds"))
} else {
  swarm_long <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_10th_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(swarm_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(width = 0.15, alpha = 0.15, size = 0.05, groupOnX = TRUE) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/tahoe_10th_paired_swarm.png"),
  width = 8, height = 5, dpi = 400)
```

## 90th Results {.tabset}
```{r}
if (do_save_tahoe) {
  tahoe_no_change_df   <- readRDS(file.path(RESULTS_PATH, "tahoe/no_change_eval_table.rds"))
  tahoe_ctrl_mean_df   <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_90th_mean_eval_table.rds"))
  tahoe_ctrl_net_df    <- readRDS(file.path(RESULTS_PATH, "tahoe/network_prop_90th_eval_table.rds"))
  tahoe_ctrl_proj_e_df <- readRDS(file.path(RESULTS_PATH, "tahoe/90th_projectcor_eigen_eval_table.rds"))
  tahoe_ctrl_proj_g_df <- readRDS(file.path(RESULTS_PATH, "tahoe/90th_projectcor_gsva_eval_table.rds"))
}
```

```{r}
if (do_save_tahoe) {
  plot_df <- bind_rows(
    tahoe_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),

    tahoe_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),

    tahoe_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    dplyr::group_by(drug, method) %>%
    dplyr::summarise_all(mean) %>%
    dplyr::ungroup() %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "tahoe/ctrl_90th_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(tahoe_ctrl_mean_df, "Mean"),
    summarize_method(tahoe_ctrl_net_df, "NetworkProp"),
    summarize_method(tahoe_ctrl_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(tahoe_ctrl_proj_g_df, "ProjectCor (GSVA)")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "tahoe/ctrl_90th_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_90th_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_90th_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +

  geom_violin(
    width = 0.9,
    alpha = 0.35,
    color = NA
  ) +

  ggbeeswarm::geom_beeswarm(
    size = 0.1,
    alpha = 0.7,
    cex = 0.7
  ) +

  stat_summary(
    fun = median,
    geom = "point",
    shape = 21,
    size = 3,
    fill = "white"
  ) +

  geom_hline(
    data = baseline_means,
    aes(yintercept = yintercept),
    linetype = "dotted",
    color = "grey50",
    linewidth = 0.8
  ) +

  facet_wrap(
    ~metric,
    scales = "free_y"
  ) +

  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)")) + 
  labs(
    x = "",
    y = "Score"
  ) +

  theme_classic(base_size = 15) +

  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/tahoe_90th_unpaired.png"),
  width = 8,
  height = 5,
  dpi = 400
)
```

### Paired {.tabset}

#### Swarm
```{r}
if (do_save_tahoe) {
  swarm_df <- bind_rows(
    tahoe_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    tahoe_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  swarm_long <- swarm_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(swarm_long, file.path(RESULTS_PATH, "tahoe/ctrl_90th_paired_swarm_plot_df.rds"))
} else {
  swarm_long <- readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_90th_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(swarm_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(width = 0.15, alpha = 0.15, size = 0.05, groupOnX = TRUE) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/tahoe_90th_paired_swarm.png"),
  width = 8, height = 5, dpi = 400)
```

# Drugmatrix {.tabset}

## Control {.tabset}
```{r}
if (do_save_drugmatrix) {
  drugmatrix_no_change_df   <- readRDS(file.path(RESULTS_PATH, "drugmatrix/no_change_eval_table.rds"))
  drugmatrix_ctrl_mean_df   <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_mean_eval_table.rds"))
  drugmatrix_ctrl_net_df    <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_networkprop_eval_table.rds"))
  drugmatrix_ctrl_proj_e_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_projectcor_eigen_eval_table.rds"))
  drugmatrix_ctrl_proj_g_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_projectcor_gsva_eval_table.rds"))
  drugmatrix_ctrl_orthos_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_orthos_residual_eval_table.rds"))
}
```

### Unpaired

```{r}
if (do_save_drugmatrix) {
  plot_df <- bind_rows(
    drugmatrix_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),

    drugmatrix_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),

    drugmatrix_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),

    drugmatrix_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),

    drugmatrix_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE),
  drugmatrix_ctrl_orthos_df %>%
      transmute(drug = gene, method = "Orthos", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "drugmatrix/ctrl_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(drugmatrix_ctrl_mean_df, "Mean"),
    summarize_method(drugmatrix_ctrl_net_df, "NetworkProp"),
    summarize_method(drugmatrix_ctrl_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(drugmatrix_ctrl_proj_g_df, "ProjectCor (GSVA)"),
    summarize_method(drugmatrix_ctrl_orthos_df, "Orthos")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "drugmatrix/ctrl_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_paired_plot_df.rds"))
}
```
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +

  geom_violin(
    width = 0.9,
    alpha = 0.35,
    color = NA
  ) +

  ggbeeswarm::geom_beeswarm(
    size = 0.1,
    alpha = 0.7,
    cex = 0.7
  ) +

  stat_summary(
    fun = median,
    geom = "point",
    shape = 21,
    size = 3,
    fill = "white"
  ) +

  geom_hline(
    data = baseline_means,
    aes(yintercept = yintercept),
    linetype = "dotted",
    color = "grey50",
    linewidth = 0.8
  ) +

  facet_wrap(
    ~metric,
    scales = "free_y"
  ) +

  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Orthos")) + 
  labs(
    x = "",
    y = "Score"
  ) +

  theme_classic(base_size = 15) +

  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )
```

```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/drugmatrix_ctrl_unpaired.png"),
  width = 8,
  height = 5,
  dpi = 400
)
```

### Paired

#### Swarm
```{r}
if (do_save_drugmatrix) {
  plot_df <- bind_rows(
    drugmatrix_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),

    drugmatrix_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),

    drugmatrix_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),

    drugmatrix_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),

    drugmatrix_ctrl_orthos_df %>%
      transmute(drug = gene, method = "Orthos", jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )

  plot_long <- plot_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "drugmatrix/ctrl_paired_swarm_plot_df.rds"))
} else {
  plot_long <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_paired_swarm_plot_df.rds"))
}
```


```{r}
ggplot(plot_long, aes(method, value, fill = method)) +

  geom_violin(
    width = 0.9,
    alpha = 0.35,
    color = NA
  ) +

  ggbeeswarm::geom_beeswarm(
    size = 0.1,
    alpha = 0.7,
    cex = 0.7
  ) +

  stat_summary(
    fun = median,
    geom = "point",
    shape = 21,
    size = 3,
    fill = "white"
  ) +

  geom_hline(
    yintercept = 0,
    linetype = "dotted",
    color = "grey50",
    linewidth = 0.8
  ) +

  facet_wrap(
    ~metric,
    scales = "free_y"
  ) +

  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Orthos")) + 

  labs(
    x = "",
    y = "Score"
  ) +

  theme_classic(base_size = 15) +

  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )
```

```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/drugmatrix_ctrl_paired_swarm.png"),
  width = 8,
  height = 5,
  dpi = 400
)
```

## 10th Results {.tabset}
```{r}
if (do_save_drugmatrix) {
  drugmatrix_no_change_df   <- readRDS(file.path(RESULTS_PATH, "drugmatrix/no_change_eval_table.rds"))
  drugmatrix_ctrl_mean_df   <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_10th_mean_eval_table.rds"))
  drugmatrix_ctrl_net_df    <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_10th_networkprop_eval_table.rds"))
  drugmatrix_ctrl_proj_e_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/10th_projectcor_eigen_eval_table.rds"))
  drugmatrix_ctrl_proj_g_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/10th_projectcor_gsva_eval_table.rds"))
}
```

```{r}
if (do_save_drugmatrix) {
  plot_df <- bind_rows(
    drugmatrix_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),

    drugmatrix_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),

    drugmatrix_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),

    drugmatrix_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),

    drugmatrix_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    dplyr::group_by(drug, method) %>%
    dplyr::summarise_all(mean) %>%
    dplyr::ungroup() %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "drugmatrix/ctrl_10th_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(drugmatrix_ctrl_mean_df, "Mean"),
    summarize_method(drugmatrix_ctrl_net_df, "NetworkProp"),
    summarize_method(drugmatrix_ctrl_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(drugmatrix_ctrl_proj_g_df, "ProjectCor (GSVA)")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "drugmatrix/ctrl_10th_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_10th_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_10th_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +

  geom_violin(
    width = 0.9,
    alpha = 0.35,
    color = NA
  ) +

  ggbeeswarm::geom_beeswarm(
    size = 0.1,
    alpha = 0.7,
    cex = 0.7
  ) +

  stat_summary(
    fun = median,
    geom = "point",
    shape = 21,
    size = 3,
    fill = "white"
  ) +

  geom_hline(
    data = baseline_means,
    aes(yintercept = yintercept),
    linetype = "dotted",
    color = "grey50",
    linewidth = 0.8
  ) +

  facet_wrap(
    ~metric,
    scales = "free_y"
  ) +

  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)")) + 
  labs(
    x = "",
    y = "Score"
  ) +

  theme_classic(base_size = 15) +

  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/drugmatrix_10th_unpaired.png"),
  width = 8,
  height = 5,
  dpi = 400
)
```

### Paired {.tabset}

#### Swarm
```{r}
if (do_save_drugmatrix) {
  swarm_df <- bind_rows(
    drugmatrix_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    drugmatrix_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    drugmatrix_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    drugmatrix_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  swarm_long <- swarm_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(swarm_long, file.path(RESULTS_PATH, "drugmatrix/ctrl_10th_paired_swarm_plot_df.rds"))
} else {
  swarm_long <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_10th_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(swarm_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.1, alpha = 0.7, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/drugmatrix_10th_paired_swarm.png"),
  width = 8, height = 5, dpi = 400)
```

## 90th Results {.tabset}
```{r}
if (do_save_drugmatrix) {
  drugmatrix_no_change_df   <- readRDS(file.path(RESULTS_PATH, "drugmatrix/no_change_eval_table.rds"))
  drugmatrix_ctrl_mean_df   <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_90th_mean_eval_table.rds"))
  drugmatrix_ctrl_net_df    <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_90th_networkprop_eval_table.rds"))
  drugmatrix_ctrl_proj_e_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/90th_projectcor_eigen_eval_table.rds"))
  drugmatrix_ctrl_proj_g_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/90th_projectcor_gsva_eval_table.rds"))
}
```

```{r}
if (do_save_drugmatrix) {
  plot_df <- bind_rows(
    drugmatrix_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),

    drugmatrix_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),

    drugmatrix_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),

    drugmatrix_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),

    drugmatrix_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    dplyr::group_by(drug, method) %>%
    dplyr::summarise_all(mean) %>%
    dplyr::ungroup() %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "drugmatrix/ctrl_90th_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(drugmatrix_ctrl_mean_df, "Mean"),
    summarize_method(drugmatrix_ctrl_net_df, "NetworkProp"),
    summarize_method(drugmatrix_ctrl_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(drugmatrix_ctrl_proj_g_df, "ProjectCor (GSVA)")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "drugmatrix/ctrl_90th_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_90th_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_90th_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +

  geom_violin(
    width = 0.9,
    alpha = 0.35,
    color = NA
  ) +

  ggbeeswarm::geom_beeswarm(
    size = 0.1,
    alpha = 0.7,
    cex = 0.7
  ) +

  stat_summary(
    fun = median,
    geom = "point",
    shape = 21,
    size = 3,
    fill = "white"
  ) +

  geom_hline(
    data = baseline_means,
    aes(yintercept = yintercept),
    linetype = "dotted",
    color = "grey50",
    linewidth = 0.8
  ) +

  facet_wrap(
    ~metric,
    scales = "free_y"
  ) +

  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)")) + 
  labs(
    x = "",
    y = "Score"
  ) +

  theme_classic(base_size = 15) +

  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/drugmatrix_90th_unpaired.png"),
  width = 8,
  height = 5,
  dpi = 400
)
```

### Paired {.tabset}

#### Swarm
```{r}
if (do_save_drugmatrix) {
  swarm_df <- bind_rows(
    drugmatrix_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    drugmatrix_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    drugmatrix_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    drugmatrix_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  swarm_long <- swarm_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(swarm_long, file.path(RESULTS_PATH, "drugmatrix/ctrl_90th_paired_swarm_plot_df.rds"))
} else {
  swarm_long <- readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_90th_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(swarm_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_beeswarm(size = 0.1, alpha = 0.7, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp", "ProjectCor (Eigen)", "ProjectCor (GSVA)")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/drugmatrix_90th_paired_swarm.png"),
  width = 8, height = 5, dpi = 400)
```

# Perturb-seq {.tabset}

## Control {.tabset}
```{r}
if (do_save_perturbseq) {
  perturbseq_no_change_df   <- readRDS(file.path(RESULTS_PATH, "perturb-seq/no_change_eval_table.rds"))
  perturbseq_ctrl_mean_df   <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_mean_eval_table.rds"))
  perturbseq_ctrl_net_df    <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_networkprop_eval_table.rds"))
  perturbseq_ctrl_proj_e_df <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_projectcor_eigen_eval_table.rds"))
  perturbseq_ctrl_proj_g_df <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_projectcor_gsva_eval_table.rds"))
  perturbseq_ctrl_stack_df  <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_stack_eval_table.rds"))
}
```

```{r}
if (do_save_perturbseq) {
  plot_df <- bind_rows(
    perturbseq_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),
    perturbseq_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_ctrl_stack_df %>%
      transmute(drug = gene, method = "Stack", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "perturb-seq/ctrl_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(perturbseq_ctrl_mean_df,   "Mean"),
    summarize_method(perturbseq_ctrl_net_df,    "NetworkProp"),
    summarize_method(perturbseq_ctrl_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(perturbseq_ctrl_proj_g_df, "ProjectCor (GSVA)"),
    summarize_method(perturbseq_ctrl_stack_df,  "Stack")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "perturb-seq/ctrl_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +
  geom_violin(width = 1.4, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.05, alpha = 0.2, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(data = baseline_means, aes(yintercept = yintercept),
             linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/perturbseq_ctrl_unpaired.png"),
       width = 8, height = 5, dpi = 400)
```

### Paired

#### Swarm
```{r}
if (do_save_perturbseq) {
  plot_df <- bind_rows(
    perturbseq_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_ctrl_stack_df %>%
      transmute(drug = gene, method = "Stack",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  plot_long <- plot_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(plot_long, file.path(RESULTS_PATH, "perturb-seq/ctrl_paired_swarm_plot_df.rds"))
} else {
  plot_long <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(plot_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.05, alpha = 0.2, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/perturbseq_ctrl_paired_swarm.png"),
       width = 8, height = 5, dpi = 400)
```


## 10th Results {.tabset}
```{r}
if (do_save_perturbseq) {
  perturbseq_no_change_df   <- readRDS(file.path(RESULTS_PATH, "perturb-seq/no_change_eval_table.rds"))
  perturbseq_10th_mean_df   <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_10th_mean_eval_table.rds"))
  perturbseq_10th_net_df    <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_10th_networkprop_eval_table.rds"))
  perturbseq_10th_proj_e_df <- readRDS(file.path(RESULTS_PATH, "perturb-seq/10th_projectcor_eigen_eval_table.rds"))
  perturbseq_10th_proj_g_df <- readRDS(file.path(RESULTS_PATH, "perturb-seq/10th_projectcor_gsva_eval_table.rds"))
  perturbseq_10th_scgpt_df  <- readRDS(file.path(RESULTS_PATH, "perturb-seq/scgpt_10th_eval_table.rds"))
  perturbseq_10th_stack_df  <- readRDS(file.path(RESULTS_PATH, "perturb-seq/10th_stack_eval_table.rds"))
}
```

```{r}
if (do_save_perturbseq) {
  plot_df <- bind_rows(
    perturbseq_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),
    perturbseq_10th_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_10th_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_10th_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_10th_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_10th_scgpt_df %>%
      transmute(drug = gene, method = "scGPT", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_10th_stack_df %>%
      transmute(drug = gene, method = "Stack", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    dplyr::group_by(drug, method) %>%
    dplyr::summarise_all(mean) %>%
    dplyr::ungroup() %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "perturb-seq/ctrl_10th_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(perturbseq_10th_mean_df,   "Mean"),
    summarize_method(perturbseq_10th_net_df,    "NetworkProp"),
    summarize_method(perturbseq_10th_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(perturbseq_10th_proj_g_df, "ProjectCor (GSVA)"),
    summarize_method(perturbseq_10th_scgpt_df,  "scGPT"),
    summarize_method(perturbseq_10th_stack_df,  "Stack")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "perturb-seq/ctrl_10th_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_10th_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_10th_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.05, alpha = 0.2, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(data = baseline_means, aes(yintercept = yintercept),
             linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "scGPT", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/perturbseq_10th_unpaired.png"),
       width = 8, height = 5, dpi = 400)
```

### Paired {.tabset}

#### Swarm
```{r}
if (do_save_perturbseq) {
  swarm_df <- bind_rows(
    perturbseq_10th_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_10th_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_10th_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_10th_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_10th_scgpt_df %>%
      transmute(drug = gene, method = "scGPT",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_10th_stack_df %>%
      transmute(drug = gene, method = "Stack",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  swarm_long <- swarm_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(swarm_long, file.path(RESULTS_PATH, "perturb-seq/ctrl_10th_paired_swarm_plot_df.rds"))
} else {
  swarm_long <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_10th_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(swarm_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.05, alpha = 0.2, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "scGPT", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/perturbseq_10th_paired_swarm.png"),
       width = 8, height = 5, dpi = 400)
```

## 90th Results {.tabset}
```{r}
if (do_save_perturbseq) {
  perturbseq_no_change_df   <- readRDS(file.path(RESULTS_PATH, "perturb-seq/no_change_eval_table.rds"))
  perturbseq_90th_mean_df   <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_90th_mean_eval_table.rds"))
  perturbseq_90th_net_df    <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_90th_networkprop_eval_table.rds"))
  perturbseq_90th_proj_e_df <- readRDS(file.path(RESULTS_PATH, "perturb-seq/90th_projectcor_eigen_eval_table.rds"))
  perturbseq_90th_proj_g_df <- readRDS(file.path(RESULTS_PATH, "perturb-seq/90th_projectcor_gsva_eval_table.rds"))
  perturbseq_90th_scgpt_df  <- readRDS(file.path(RESULTS_PATH, "perturb-seq/scgpt_90th_eval_table.rds"))
  perturbseq_90th_stack_df  <- readRDS(file.path(RESULTS_PATH, "perturb-seq/90th_stack_eval_table.rds"))
}
```

```{r}
if (do_save_perturbseq) {
  plot_df <- bind_rows(
    perturbseq_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),
    perturbseq_90th_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_90th_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_90th_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_90th_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_90th_scgpt_df %>%
      transmute(drug = gene, method = "scGPT", jacc = jacc_TRUE, NES = NES_TRUE),
    perturbseq_90th_stack_df %>%
      transmute(drug = gene, method = "Stack", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    dplyr::group_by(drug, method) %>%
    dplyr::summarise_all(mean) %>%
    dplyr::ungroup() %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "perturb-seq/ctrl_90th_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(perturbseq_90th_mean_df,   "Mean"),
    summarize_method(perturbseq_90th_net_df,    "NetworkProp"),
    summarize_method(perturbseq_90th_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(perturbseq_90th_proj_g_df, "ProjectCor (GSVA)"),
    summarize_method(perturbseq_90th_scgpt_df,  "scGPT"),
    summarize_method(perturbseq_90th_stack_df,  "Stack")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "perturb-seq/ctrl_90th_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_90th_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_90th_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.05, alpha = 0.2, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(data = baseline_means, aes(yintercept = yintercept),
             linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "scGPT", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/perturbseq_90th_unpaired.png"),
       width = 8, height = 5, dpi = 400)
```

### Paired {.tabset}

#### Swarm
```{r}
if (do_save_perturbseq) {
  swarm_df <- bind_rows(
    perturbseq_90th_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_90th_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_90th_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_90th_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_90th_scgpt_df %>%
      transmute(drug = gene, method = "scGPT",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    perturbseq_90th_stack_df %>%
      transmute(drug = gene, method = "Stack",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  swarm_long <- swarm_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(swarm_long, file.path(RESULTS_PATH, "perturb-seq/ctrl_90th_paired_swarm_plot_df.rds"))
} else {
  swarm_long <- readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_90th_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(swarm_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.05, alpha = 0.2, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "scGPT", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/perturbseq_90th_paired_swarm.png"),
       width = 8, height = 5, dpi = 400)
```

# SciPlex {.tabset}

## Control {.tabset}
```{r}
if (do_save_sciplex) {
  sciplex_no_change_df   <- readRDS(file.path(RESULTS_PATH, "sciplex/no_change_eval_table.rds"))
  sciplex_ctrl_mean_df   <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_mean_eval_table.rds"))
  sciplex_ctrl_net_df    <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_networkprop_eval_table.rds"))
  sciplex_ctrl_proj_e_df <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_projectcor_eigen_eval_table.rds"))
  sciplex_ctrl_proj_g_df <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_projectcor_gsva_eval_table.rds"))
  sciplex_ctrl_orthos_df <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_orthos_residual_eval_table.rds"))
  sciplex_ctrl_stack_df  <- readRDS(file.path(RESULTS_PATH, "sciplex/stack_ctrl_eval_table.rds"))
}
```

```{r}
if (do_save_sciplex) {
  plot_df <- bind_rows(
    sciplex_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),
    sciplex_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_ctrl_orthos_df %>%
      transmute(drug = gene, method = "Orthos", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_ctrl_stack_df %>%
      transmute(drug = gene, method = "Stack", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "sciplex/ctrl_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(sciplex_ctrl_mean_df,   "Mean"),
    summarize_method(sciplex_ctrl_net_df,    "NetworkProp"),
    summarize_method(sciplex_ctrl_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(sciplex_ctrl_proj_g_df, "ProjectCor (GSVA)"),
    summarize_method(sciplex_ctrl_orthos_df, "Orthos"),
    summarize_method(sciplex_ctrl_stack_df,  "Stack")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "sciplex/ctrl_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_beeswarm(size = 0.1, alpha = 0.7, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(data = baseline_means, aes(yintercept = yintercept),
             linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Orthos", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/sciplex_ctrl_unpaired.png"),
       width = 8, height = 5, dpi = 400)
```

### Paired

#### Swarm
```{r}
if (do_save_sciplex) {
  plot_df <- bind_rows(
    sciplex_ctrl_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_ctrl_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_ctrl_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_ctrl_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_ctrl_orthos_df %>%
      transmute(drug = gene, method = "Orthos",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_ctrl_stack_df %>%
      transmute(drug = gene, method = "Stack",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  plot_long <- plot_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(plot_long, file.path(RESULTS_PATH, "sciplex/ctrl_paired_swarm_plot_df.rds"))
} else {
  plot_long <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(plot_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_beeswarm(size = 0.1, alpha = 0.7, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Orthos", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/sciplex_ctrl_paired_swarm.png"),
       width = 8, height = 5, dpi = 400)
```


## 10th Results {.tabset}
```{r}
if (do_save_sciplex) {
  sciplex_no_change_df   <- readRDS(file.path(RESULTS_PATH, "sciplex/no_change_eval_table.rds"))
  sciplex_10th_mean_df   <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_10th_mean_eval_table.rds"))
  sciplex_10th_net_df    <- readRDS(file.path(RESULTS_PATH, "sciplex/10th_networkprop_eval_table.rds"))
  sciplex_10th_proj_e_df <- readRDS(file.path(RESULTS_PATH, "sciplex/10th_projectcor_eigen_eval_table.rds"))
  sciplex_10th_proj_g_df <- readRDS(file.path(RESULTS_PATH, "sciplex/10th_projectcor_gsva_eval_table.rds"))
  sciplex_10th_stack_df  <- readRDS(file.path(RESULTS_PATH, "sciplex/stack_10th_eval_table.rds"))
}
```

```{r}
if (do_save_sciplex) {
  plot_df <- bind_rows(
    sciplex_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),
    sciplex_10th_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_10th_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_10th_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_10th_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_10th_stack_df %>%
      transmute(drug = gene, method = "Stack", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    dplyr::group_by(drug, method) %>%
    dplyr::summarise_all(mean) %>%
    dplyr::ungroup() %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "sciplex/ctrl_10th_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(sciplex_10th_mean_df,   "Mean"),
    summarize_method(sciplex_10th_net_df,    "NetworkProp"),
    summarize_method(sciplex_10th_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(sciplex_10th_proj_g_df, "ProjectCor (GSVA)"),
    summarize_method(sciplex_10th_stack_df,  "Stack")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "sciplex/ctrl_10th_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_10th_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_10th_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.8, alpha = 0.7, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(data = baseline_means, aes(yintercept = yintercept),
             linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/sciplex_10th_unpaired.png"),
       width = 8, height = 5, dpi = 400)
```

### Paired {.tabset}

#### Swarm
```{r}
if (do_save_sciplex) {
  swarm_df <- bind_rows(
    sciplex_10th_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_10th_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_10th_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_10th_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_10th_stack_df %>%
      transmute(drug = gene, method = "Stack",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  swarm_long <- swarm_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(swarm_long, file.path(RESULTS_PATH, "sciplex/ctrl_10th_paired_swarm_plot_df.rds"))
} else {
  swarm_long <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_10th_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(swarm_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.1, alpha = 0.7, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/sciplex_10th_paired_swarm.png"),
       width = 8, height = 5, dpi = 400)
```

## 90th Results {.tabset}
```{r}
if (do_save_sciplex) {
  sciplex_no_change_df   <- readRDS(file.path(RESULTS_PATH, "sciplex/no_change_eval_table.rds"))
  sciplex_90th_mean_df   <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_90th_mean_eval_table.rds"))
  sciplex_90th_net_df    <- readRDS(file.path(RESULTS_PATH, "sciplex/90th_networkprop_eval_table.rds"))
  sciplex_90th_proj_e_df <- readRDS(file.path(RESULTS_PATH, "sciplex/90th_projectcor_eigen_eval_table.rds"))
  sciplex_90th_proj_g_df <- readRDS(file.path(RESULTS_PATH, "sciplex/90th_projectcor_gsva_eval_table.rds"))
  sciplex_90th_stack_df  <- readRDS(file.path(RESULTS_PATH, "sciplex/stack_90th_eval_table.rds"))
}
```

```{r}
if (do_save_sciplex) {
  plot_df <- bind_rows(
    sciplex_no_change_df %>%
      transmute(drug = gene, method = "No change", jacc, NES),
    sciplex_90th_mean_df %>%
      transmute(drug = gene, method = "Mean", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_90th_net_df %>%
      transmute(drug = gene, method = "NetworkProp", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_90th_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_90th_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)", jacc = jacc_TRUE, NES = NES_TRUE),
    sciplex_90th_stack_df %>%
      transmute(drug = gene, method = "Stack", jacc = jacc_TRUE, NES = NES_TRUE)
  )

  plot_long <- plot_df %>%
    dplyr::rename(Jaccard = jacc) %>%
    dplyr::group_by(drug, method) %>%
    dplyr::summarise_all(mean) %>%
    dplyr::ungroup() %>%
    pivot_longer(c(Jaccard, NES), names_to = "metric", values_to = "value")

  saveRDS(plot_long, file.path(RESULTS_PATH, "sciplex/ctrl_90th_unpaired_plot_df.rds"))

  summary_df <- bind_rows(
    summarize_method(sciplex_90th_mean_df,   "Mean"),
    summarize_method(sciplex_90th_net_df,    "NetworkProp"),
    summarize_method(sciplex_90th_proj_e_df, "ProjectCor (Eigen)"),
    summarize_method(sciplex_90th_proj_g_df, "ProjectCor (GSVA)"),
    summarize_method(sciplex_90th_stack_df,  "Stack")
  )
  saveRDS(summary_df, file.path(RESULTS_PATH, "sciplex/ctrl_90th_paired_plot_df.rds"))
} else {
  plot_long  <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_90th_unpaired_plot_df.rds"))
  summary_df <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_90th_paired_plot_df.rds"))
}
```

### Unpaired
```{r}
baseline_means <- plot_long %>%
  filter(method == "No change") %>%
  group_by(metric) %>%
  summarise(yintercept = median(value), .groups = "drop")

ggplot(plot_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_quasirandom(size = 0.8, alpha = 0.7, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(data = baseline_means, aes(yintercept = yintercept),
             linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("No change", "Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/sciplex_90th_unpaired.png"),
       width = 8, height = 5, dpi = 400)
```

### Paired {.tabset}

#### Swarm
```{r}
if (do_save_sciplex) {
  swarm_df <- bind_rows(
    sciplex_90th_mean_df %>%
      transmute(drug = gene, method = "Mean",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_90th_net_df %>%
      transmute(drug = gene, method = "NetworkProp",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_90th_proj_e_df %>%
      transmute(drug = gene, method = "ProjectCor (Eigen)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_90th_proj_g_df %>%
      transmute(drug = gene, method = "ProjectCor (GSVA)",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE),
    sciplex_90th_stack_df %>%
      transmute(drug = gene, method = "Stack",
                jacc_delta = jacc_TRUE - jacc_FALSE, NES_delta = NES_TRUE - NES_FALSE)
  )
  swarm_long <- swarm_df %>%
    rename(`Jaccard ∆` = jacc_delta, `NES ∆` = NES_delta) %>%
    pivot_longer(c(`Jaccard ∆`, `NES ∆`), names_to = "metric", values_to = "value")
  saveRDS(swarm_long, file.path(RESULTS_PATH, "sciplex/ctrl_90th_paired_swarm_plot_df.rds"))
} else {
  swarm_long <- readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_90th_paired_swarm_plot_df.rds"))
}
```
```{r}
ggplot(swarm_long, aes(method, value, fill = method)) +
  geom_violin(width = 0.9, alpha = 0.35, color = NA) +
  ggbeeswarm::geom_beeswarm(size = 0.1, alpha = 0.7, cex = 0.7) +
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50", linewidth = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_x_discrete(limits = c("Mean", "NetworkProp",
                               "ProjectCor (Eigen)", "ProjectCor (GSVA)", "Stack")) +
  labs(x = "", y = "Score") +
  theme_classic(base_size = 15) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "none",
        strip.text = element_text(face = "bold"))
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/sciplex_90th_paired_swarm.png"),
       width = 8, height = 5, dpi = 400)
```

# Tahoe MOA {.tabset}

## Delta NES Across Methods
```{r}
tahoe_metadata_path <- file.path(
  Sys.getenv("AGED"),
  "CBMrepositoryData/perturbational_data/tahoe/metadata/drug_metadata.parquet"
)

normalize_drug_name <- function(x) {
  x %>%
    str_to_lower() %>%
    str_replace_all("[^[:alnum:]]+", "")
}

normalize_stack_target <- function(source, target) {
  if_else(
    target == "target_context" & str_detect(source, "^nci_h23_"),
    str_remove(source, "^nci_h23_"),
    target
  )
}

clean_moa_fine_label <- function(x) {
  x %>%
    str_remove(regex("\\s+(agonist|activator|inhibitor|antagonist)$", ignore_case = TRUE)) %>%
    str_squish()
}

clean_moa_broad_label <- function(x) {
  recode(
    x,
    "activator/agonist" = "Activator/\nAgonist",
    "inhibitor/antagonist" = "Inhibitor/\nAntagonist"
  )
}

if (do_save_tahoe) {
  tahoe_drug_metadata <- arrow::read_parquet(tahoe_metadata_path) %>%
    as_tibble() %>%
    select(drug, `moa-broad`, `moa-fine`) %>%
    distinct() %>%
    mutate(drug_key = normalize_drug_name(drug)) %>%
    distinct(drug_key, .keep_all = TRUE)

  tahoe_eval_specs <- tribble(
    ~method, ~regime, ~file,
    "Mean", "ctrl", "ctrl_mean_eval_table.rds",
    "Mean", "10th", "ctrl_10th_mean_eval_table.rds",
    "Mean", "90th", "ctrl_90th_mean_eval_table.rds",
    "NetworkProp", "ctrl", "ctrl_networkprop_eval_table.rds",
    "NetworkProp", "10th", "network_prop_10th_eval_table.rds",
    "NetworkProp", "90th", "network_prop_90th_eval_table.rds",
    "ProjectCor (Eigen)", "ctrl", "ctrl_projectcor_eigen_eval_table.rds",
    "ProjectCor (Eigen)", "10th", "10th_projectcor_eigen_eval_table.rds",
    "ProjectCor (Eigen)", "90th", "90th_projectcor_eigen_eval_table.rds",
    "ProjectCor (GSVA)", "ctrl", "ctrl_projectcor_gsva_eval_table.rds",
    "ProjectCor (GSVA)", "10th", "10th_projectcor_gsva_eval_table.rds",
    "ProjectCor (GSVA)", "90th", "90th_projectcor_gsva_eval_table.rds",
    "Orthos", "ctrl", "ctrl_orthos_residual_eval_table.rds",
    "Stack", "ctrl", "stack_ctrl_eval_table.rds",
    "Stack", "10th", "stack_10th_eval_table.rds",
    "Stack", "90th", "stack_90th_eval_table.rds"
  ) %>%
    mutate(path = file.path(RESULTS_PATH, "tahoe", file)) %>%
    filter(file.exists(path))

  read_tahoe_eval <- function(path, method, regime) {
    print(path)
    readRDS(path) %>%
      as_tibble() %>%
      transmute(
        method = method,
        regime = regime,
        gene,
        target = normalize_stack_target(source, target),
        delta_NES = NES_TRUE - NES_FALSE
      ) %>%
      group_by(method, regime, gene, target) %>%
      summarise(delta_NES = median(delta_NES, na.rm = TRUE), .groups = "drop")
  }

  tahoe_moa_delta_joined_df <- pmap_dfr(
    tahoe_eval_specs %>% select(path, method, regime),
    read_tahoe_eval
  ) %>%
    mutate(regime = if_else(regime == "ctrl", "Control", regime)) %>%
    mutate(
      method = factor(
        method,
        levels = c("Mean", "NetworkProp", "ProjectCor (Eigen)",
                   "ProjectCor (GSVA)", "Orthos", "Stack")
      ),
      regime = factor(regime, levels = c("90th", "10th", "Control")),
      drug_key = normalize_drug_name(gene)
    ) %>%
    left_join(tahoe_drug_metadata, by = "drug_key")

  tahoe_unmatched_moa_drugs <- tahoe_moa_delta_joined_df %>%
    filter(is.na(`moa-fine`)) %>%
    distinct(gene) %>%
    arrange(gene)

  saveRDS(
    tahoe_unmatched_moa_drugs,
    file.path(RESULTS_PATH, "tahoe/tahoe_moa_delta_unmatched_drugs.rds")
  )

  tahoe_moa_delta_df <- tahoe_moa_delta_joined_df %>%
    filter(
      !is.na(`moa-fine`),
      `moa-fine` != "unclear",
      `moa-broad` %in% c("activator/agonist", "inhibitor/antagonist")
    ) %>%
    mutate(
      moa_broad_group = factor(
        clean_moa_broad_label(`moa-broad`),
        levels = c("Inhibitor/\nAntagonist", "Activator/\nAgonist")
      ),
      moa_fine_label = clean_moa_fine_label(`moa-fine`)
    )

  saveRDS(
    tahoe_moa_delta_df,
    file.path(RESULTS_PATH, "tahoe/tahoe_moa_delta_nes_plot_df.rds")
  )
} else {
  tahoe_unmatched_moa_drugs <- readRDS(file.path(RESULTS_PATH, "tahoe/tahoe_moa_delta_unmatched_drugs.rds"))
  tahoe_moa_delta_df        <- readRDS(file.path(RESULTS_PATH, "tahoe/tahoe_moa_delta_nes_plot_df.rds"))
}
```

```{r, fig.width=14, fig.height=10}
moa_order <- tahoe_moa_delta_df %>%
  filter(regime == "90th") %>%
  group_by(moa_broad_group, moa_fine_label) %>%
  summarise(median_delta_NES = median(delta_NES, na.rm = TRUE), .groups = "drop") %>%
  group_by(moa_broad_group) %>%
  arrange(desc(median_delta_NES)) %>%
  summarise(moa_fine_label = list(moa_fine_label), .groups = "drop")

moa_levels <- c(
  rev(moa_order$moa_fine_label[[which(moa_order$moa_broad_group == "Activator/\nAgonist")]]),
  rev(moa_order$moa_fine_label[[which(moa_order$moa_broad_group == "Inhibitor/\nAntagonist")]])
)

moa_label_counts <- tahoe_moa_delta_df %>%
  group_by(moa_fine_label) %>%
  summarise(n_drugs = n_distinct(gene), .groups = "drop")

format_moa_axis_label <- function(x) {
  n_drugs <- moa_label_counts$n_drugs[match(x, moa_label_counts$moa_fine_label)]
  str_wrap(paste0(x, " (", n_drugs, ")"), width = 28)
}

p_moa_delta <- ggplot(
  tahoe_moa_delta_df %>%
    mutate(
      moa_fine_label = factor(moa_fine_label, levels = moa_levels),
      method = dplyr::recode(as.character(method),
        "NetworkProp"        = "NetProp",
        "ProjectCor (Eigen)" = "projCor-Eigen",
        "ProjectCor (GSVA)"  = "projCor-GSVA"
      )
    ),
  aes(x = delta_NES, y = moa_fine_label)
) +
  geom_vline(xintercept = 0, color = "grey70", linewidth = 0.35) +
  geom_boxplot(
    aes(fill = regime),
    outlier.shape = NA,
    alpha = 0.65,
    width = 0.55
  ) +
  ggbeeswarm::geom_quasirandom(
    aes(color = method, group = moa_fine_label),
    width = 0.18,
    size = 0.6,
    alpha = 0.25,
    groupOnX = FALSE
  ) +
  facet_grid(
    rows = vars(moa_broad_group),
    cols = vars(regime),
    scales = "free_y",
    space = "free_y",
    switch = "y"
  ) +
  scale_fill_brewer(palette = "Pastel1", guide = "none") +
  scale_y_discrete(labels = format_moa_axis_label, position = "right") +
  labs(
    x = "Delta NES",
    y = NULL,
    color = "Method"
  ) +
  theme_bw(base_size = 9) +
  theme(
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    strip.text.y.left = element_text(face = "bold", angle = 0),
    legend.position = "right",
  )
```

```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/paper_fig4_tahoe_moa.pdf"),
       p_moa_delta,
       width = 8,
       height = 5.17,
       dpi = 400,
       device = "pdf")
```

## Gain Heatmaps
```{r}
if (do_save_tahoe) {
  tahoe_moa_gain_df <- tahoe_moa_delta_df %>%
    select(method, regime, gene, target, moa_broad_group, moa_fine_label, delta_NES) %>%
    pivot_wider(names_from = regime, values_from = delta_NES) %>%
    mutate(
      gain_10th = `10th` - Control,
      gain_90th = `90th` - Control
    ) %>%
    pivot_longer(
      c(gain_10th, gain_90th),
      names_to = "comparison",
      values_to = "gain"
    ) %>%
    filter(is.finite(gain)) %>%
    group_by(comparison, moa_broad_group, moa_fine_label, method) %>%
    summarise(median_gain = median(gain, na.rm = TRUE), .groups = "drop")

  saveRDS(
    tahoe_moa_gain_df,
    file.path(RESULTS_PATH, "tahoe/tahoe_moa_gain_heatmap_df.rds")
  )
} else {
  tahoe_moa_gain_df <- readRDS(file.path(RESULTS_PATH, "tahoe/tahoe_moa_gain_heatmap_df.rds"))
}

plot_gain_heatmap <- function(gain_df, comparison_name, title) {
  gain_wide <- gain_df %>%
    filter(comparison == comparison_name) %>%
    select(moa_broad_group, moa_fine_label, method, median_gain) %>%
    pivot_wider(names_from = method, values_from = median_gain) %>%
    arrange(moa_broad_group, moa_fine_label)
  
  row_group <- gain_wide$moa_broad_group
  row_group <- factor(row_group, levels = c("Inhibitor/\nAntagonist", "Activator/\nAgonist"))
  gain_mat <- gain_wide %>%
    select(-moa_broad_group) %>%
    column_to_rownames("moa_fine_label") %>%
    as.matrix()
  
  max_abs <- max(abs(gain_mat), na.rm = TRUE)
  
  Heatmap(
    gain_mat,
    name = "Median gain",
    col = circlize::colorRamp2(c(-max_abs, 0, max_abs), c("#3b4cc0", "white", "#b40426")),
    na_col = "grey90",
    cluster_rows = TRUE,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    cluster_columns = TRUE,
    row_split = row_group,
    row_gap = unit(5, "mm"),
    cluster_row_slices = FALSE,
    row_names_side = "right",
    row_names_gp = gpar(fontsize = 17), 
    column_names_rot = 45,
    column_names_gp = gpar(fontsize = 17), 
    column_title = title,
    column_title_gp = gpar(fontsize = 17, fontface = "bold"), 
    row_title_side = "left",
    row_title_rot = 0,
    row_title_gp = gpar(fontsize = 17, fontface = "bold"),
    cell_fun = function(j, i, x, y, width, height, fill) {
      value <- gain_mat[i, j]
      if (is.finite(value)) {
        grid::grid.text(sprintf("%.2f", value), x, y, gp = grid::gpar(fontsize = 14))
      }
    },
    heatmap_legend_param = list(title = "Median\nDelta NES",
                                title_gp = gpar(fontsize = 16, fontface = "bold"),
                                labels_gp = gpar(fontsize = 15, col = "black"))
  )
}

ht_10th <- plot_gain_heatmap(tahoe_moa_gain_df, "gain_10th", "10th - ctrl")
ht_90th <- plot_gain_heatmap(tahoe_moa_gain_df, "gain_90th", "90th - ctrl")
```

```{r, fig.width=11, fig.height=10}
draw(ht_10th)
```

```{r}
png(file.path(PROJECT_PATH, "results/plots/tahoe_moa_gain_10th_minus_ctrl_heatmap.png"),
    width = 11, height = 10, units = "in", res = 400)
draw(ht_10th)
dev.off()
```

```{r, fig.width=11, fig.height=10}
draw(ht_90th)
```

```{r}
png(file.path(PROJECT_PATH, "results/plots/tahoe_moa_gain_90th_minus_ctrl_heatmap.png"),
    width = 11, height = 10, units = "in", res = 400)
draw(ht_90th)
dev.off()
```

## Gain Lineplots
```{r}
tahoe_moa_regime_df <- tahoe_moa_delta_df %>%
  group_by(method, regime, moa_broad_group, moa_fine_label) %>%
  summarise(median_delta_NES = median(delta_NES, na.rm = TRUE), .groups = "drop") %>%
  mutate(regime = factor(regime, levels = c("Control", "10th", "90th")))

# Rank top/bottom 3 drug classes per method based on 90th - ctrl gain
moa_gain_rank <- tahoe_moa_regime_df %>%
  pivot_wider(names_from = regime, values_from = median_delta_NES) %>%
  mutate(gain_90th = `90th` - Control) %>%
  group_by(method) %>%
  arrange(method, desc(gain_90th)) %>%
  mutate(rank = row_number(), n = n()) %>%
  mutate(highlight = case_when(
    rank <= 3 ~ "top3",
    rank > n - 3 ~ "bottom3",
    TRUE ~ "other"
  )) %>%
  ungroup() %>%
  select(method, moa_fine_label, highlight)

tahoe_moa_regime_plot_df <- tahoe_moa_regime_df %>%
  left_join(moa_gain_rank, by = c("method", "moa_fine_label"))
```

```{r, fig.width=14, fig.height=10}
tahoe_moa_gain_long_df <- tahoe_moa_regime_df %>%
  pivot_wider(names_from = regime, values_from = median_delta_NES) %>%
  mutate(
    gain_10th = `10th` - Control,
    gain_90th = `90th` - Control
  ) %>%
  pivot_longer(
    c(gain_10th, gain_90th),
    names_to = "comparison",
    values_to = "gain"
  ) %>%
  mutate(
    x_label = recode(comparison, gain_10th = "10th", gain_90th = "90th"),
    x_label = factor(x_label, levels = c("10th", "90th"))
  ) %>%
  bind_rows(
    distinct(., method, moa_broad_group, moa_fine_label) %>%
      mutate(x_label = factor("ctrl", levels = c("ctrl", "10th", "90th")), gain = 0)
  ) %>%
  mutate(x_label = factor(x_label, levels = c("ctrl", "10th", "90th")))

# Reuse same highlight assignments from Version 1
tahoe_moa_gain_long_df <- tahoe_moa_gain_long_df %>%
  left_join(moa_gain_rank, by = c("method", "moa_fine_label"))

highlight_labels_v2 <- tahoe_moa_gain_long_df %>%
  filter(highlight != "other", x_label == "90th") %>%
  select(method, moa_fine_label, x_label, gain, highlight)

ggplot(tahoe_moa_gain_long_df,
       aes(x = x_label, y = gain, group = moa_fine_label)) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50") +
  geom_line(data = filter(tahoe_moa_gain_long_df, highlight == "other"),
            color = "grey80", linewidth = 0.4, alpha = 0.7) +
  geom_line(data = filter(tahoe_moa_gain_long_df, highlight == "top3"),
            aes(color = moa_fine_label), linewidth = 1.2) +
  geom_line(data = filter(tahoe_moa_gain_long_df, highlight == "bottom3"),
            aes(color = moa_fine_label), linewidth = 1.2, linetype = "dashed") +
  geom_point(data = filter(tahoe_moa_gain_long_df, highlight != "other"),
             aes(color = moa_fine_label), size = 2.5) +
  ggrepel::geom_text_repel(
    data = highlight_labels_v2,
    aes(label = moa_fine_label, color = moa_fine_label),
    size = 3, nudge_x = 0.15, direction = "y", hjust = 0,
    segment.size = 0.3, show.legend = FALSE
  ) +
  facet_wrap(~ method, scales = "free_y") +
  scale_x_discrete(expand = expansion(mult = c(0.05, 0.35))) +
  labs(
    title = "Gain in delta NES relative to ctrl — top/bottom 3 drug classes per method",
    subtitle = "Solid = top 3 improving (90th − ctrl), dashed = bottom 3 worsening; all lines anchored at 0 = ctrl",
    x = "Regime", y = "Gain (delta NES vs. ctrl)", color = "Drug class"
  ) +
  theme_bw(base_size = 13) +
  theme(legend.position = "none", strip.text = element_text(face = "bold"))
```

```{r}
ggsave(
  file.path(PROJECT_PATH, "results/plots/tahoe_moa_gain_lineplot_relative_to_ctrl.png"),
  width = 14, height = 10, dpi = 300
)
```

## Combined Delta NES + Gain Heatmap

```{r tahoe-moa-combined, message=FALSE, warning=FALSE, fig.width=14, fig.height=20}
# Fixed method column order matching the delta NES plot legend
method_order_hm <- c("Mean", "NetProp", "projCor-Eigen", "projCor-GSVA", "Stack")

# Row order top-to-bottom in heatmap = rev(moa_levels), so that
# top of heatmap aligns with top facet (Inhibitor) of the ggplot above.
row_order_hm <- rev(moa_levels)

# Lookup for row split: moa_fine_label -> moa_broad_group
moa_to_broad <- tahoe_moa_gain_df %>%
  dplyr::distinct(moa_fine_label, moa_broad_group) %>%
  tibble::deframe()

row_split_hm <- factor(
  as.character(moa_to_broad[row_order_hm]),
  levels = c("Inhibitor/\nAntagonist", "Activator/\nAgonist")  # top slice first
)

# Build gain matrices (rows = MOA in heatmap order, cols = methods in fixed order)
make_gain_mat <- function(comp) {
  tahoe_moa_gain_df %>%
    dplyr::filter(comparison == comp) %>%
    dplyr::select(moa_fine_label, method, median_gain) %>%
    dplyr::mutate(
      moa_fine_label = as.character(moa_fine_label),
      method = dplyr::recode(as.character(method),
        "NetworkProp"        = "NetProp",
        "ProjectCor (Eigen)" = "projCor-Eigen",
        "ProjectCor (GSVA)"  = "projCor-GSVA")
    ) %>%
    tidyr::pivot_wider(names_from = method, values_from = median_gain) %>%
    tibble::column_to_rownames("moa_fine_label") %>%
    .[row_order_hm, method_order_hm] %>%
    as.matrix()
}

gain_mat_10th <- make_gain_mat("gain_10th")
gain_mat_90th <- make_gain_mat("gain_90th")

# Shared symmetric color scale across both regimes
max_abs <- max(abs(c(gain_mat_10th, gain_mat_90th)), na.rm = TRUE)
gain_col <- circlize::colorRamp2(c(-max_abs, 0, max_abs),
                                  c("#3b4cc0", "white", "#b40426"))

common_hm_args <- list(
  col            = gain_col,
  na_col         = "grey90",
  cluster_rows   = FALSE,
  cluster_columns = FALSE,
  show_row_dend  = FALSE,
  show_column_dend = FALSE,
  cluster_row_slices = FALSE,
  row_split      = row_split_hm,
  row_gap        = unit(5, "mm"),
  row_names_side = "right",
  row_names_gp   = gpar(fontsize = 14),
  column_names_rot = 45,
  column_names_gp  = gpar(fontsize = 14),
  column_title_gp  = gpar(fontsize = 15, fontface = "bold"),
  row_title_side   = "left",
  row_title_rot    = 0,
  row_title_gp     = gpar(fontsize = 14, fontface = "bold"),
  cell_fun = function(j, i, x, y, width, height, fill) {
    value <- gain_mat_10th[i, j]   # placeholder; overridden per heatmap below
    if (is.finite(value))
      grid::grid.text(sprintf("%.2f", value), x, y, gp = grid::gpar(fontsize = 11))
  },
  heatmap_legend_param = list(
    title      = "Median\nGain",
    title_gp   = gpar(fontsize = 14, fontface = "bold"),
    labels_gp  = gpar(fontsize = 13)
  )
)

ht_10th_combined <- Heatmap(
  gain_mat_10th,
  name           = "Median\nGain",
  column_title   = "10th - Control",
  cell_fun       = function(j, i, x, y, width, height, fill) {
    v <- gain_mat_10th[i, j]
    if (is.finite(v))
      grid::grid.text(sprintf("%.2f", v), x, y, gp = grid::gpar(fontsize = 8))
  },
  col              = gain_col,
  na_col           = "grey90",
  cluster_rows     = FALSE,
  cluster_columns  = FALSE,
  show_row_dend    = FALSE,
  show_column_dend = FALSE,
  cluster_row_slices = FALSE,
  row_split        = row_split_hm,
  row_gap          = unit(5, "mm"),
  row_names_side   = "right",
  row_names_gp     = gpar(fontsize = 11),
  column_names_rot = 45,
  column_names_gp  = gpar(fontsize = 11),
  column_title_gp  = gpar(fontsize = 12, fontface = "bold"),
  row_title_side   = "left",
  row_title_rot    = 0,
  row_title_gp     = gpar(fontsize = 11, fontface = "bold"),
  heatmap_legend_param = list(
    title    = "Median\nGain",
    title_gp = gpar(fontsize = 11, fontface = "bold"),
    labels_gp = gpar(fontsize = 10)
  )
)

ht_90th_combined <- Heatmap(
  gain_mat_90th,
  name           = "Median\nGain_90th",
  column_title   = "90th - Control",
  show_heatmap_legend = FALSE,
  cell_fun       = function(j, i, x, y, width, height, fill) {
    v <- gain_mat_90th[i, j]
    if (is.finite(v))
      grid::grid.text(sprintf("%.2f", v), x, y, gp = grid::gpar(fontsize = 8))
  },
  col              = gain_col,
  na_col           = "grey90",
  cluster_rows     = FALSE,
  cluster_columns  = FALSE,
  show_row_dend    = FALSE,
  show_column_dend = FALSE,
  cluster_row_slices = FALSE,
  row_names_side   = "right",
  row_names_gp     = gpar(fontsize = 11),
  column_names_rot = 45,
  column_names_gp  = gpar(fontsize = 11),
  column_title_gp  = gpar(fontsize = 12, fontface = "bold")
)

ht_combined <- ht_10th_combined + ht_90th_combined

ComplexHeatmap::draw(ht_combined, merge_legend = TRUE)
```

```{r}
pdf(file.path(PROJECT_PATH, "results/plots/tahoe_moa_gain_combined.pdf"),
    width = 8, height = 8)
ComplexHeatmap::draw(ht_combined, merge_legend = TRUE)
dev.off()
```

## TI Quadrant Scatter
```{r}
tahoe_trade_metrics <- readRDS(file.path(PROJECT_PATH,"results/plots/trade_metric_correlations/tahoe/context_trade_metrics.rds"))

tahoe_source_ti <- tahoe_trade_metrics %>%
  filter(context_key == normalize_drug_name("NCI-H23")) %>%
  mutate(drug_key = normalize_drug_name(drug)) %>%
  select(drug_key, source_TI = TI) %>%
  distinct(drug_key, .keep_all = TRUE)

tahoe_target_ti <- tahoe_trade_metrics %>%
  mutate(drug_key = normalize_drug_name(drug)) %>%
  select(target_key = context_key, drug_key, target_TI = TI) %>%
  distinct(target_key, drug_key, .keep_all = TRUE)

target_keys_to_keep <- tahoe_moa_delta_df %>%
  distinct(target_key = normalize_drug_name(target)) %>%
  pull(target_key)
drug_keys_to_keep <- tahoe_moa_delta_df %>%
  distinct(drug_key) %>% 
  pull(drug_key)
```

```{r}
source_target_ti_df <- tahoe_target_ti %>%
  left_join(tahoe_source_ti, by = "drug_key") %>%
  filter(
    is.finite(source_TI),
    is.finite(target_TI),
    target_key != normalize_drug_name("NCI-H23"),
    target_key %in% target_keys_to_keep,
    drug_key %in% drug_keys_to_keep
  )

ggplot(source_target_ti_df, aes(x = source_TI, y = target_TI)) +
  geom_abline(slope = 1, intercept = 0, color = "grey70", linewidth = 0.5, linetype = "dashed") +
  geom_point(aes(fill = target_key), shape = 21, color = "grey20", alpha = 0.75, size = 2.2) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    x = "Source TI (NCI-H23)",
    y = "Target TI",
    fill = "Target cell line"
  ) +
  theme_bw(base_size = 16) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

```


```{r}
ggsave(
  file.path(PROJECT_PATH, "results/plots/tahoe_source_target_TI_correlation.png"),
  width = 10,
  height = 10,
  dpi = 400
)
```
```{r}
drug_order <- source_target_ti_df %>%
  group_by(drug_key) %>%
  summarise(median_target_TI = median(target_TI, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(median_target_TI)) %>%
  summarise(drug_key = list(drug_key), .groups = "drop")

drug_levels <- drug_order$drug_key[[1]]
ggplot(source_target_ti_df %>%
    mutate(drug_key = factor(drug_key, levels = drug_levels)), aes(x = source_TI, y = target_TI)) +
  geom_abline(slope = 1, intercept = 0, color = "grey70", linewidth = 0.5, linetype = "dashed") +
  geom_point(aes(fill = drug_key), shape = 21, color = "grey20", alpha = 0.75, size = 2.2) +
  # scale_fill_brewer(palette = "Set2") +
  labs(
    x = "Source TI (NCI-H23)",
    y = "Target TI",
    fill = "Drug"
  ) +
  theme_bw(base_size = 16) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "none"
  )
```

```{r}
if (do_save_tahoe) {
  tahoe_moa_ti_df <- tahoe_moa_delta_df %>%
    mutate(target_key = normalize_drug_name(target)) %>%
    left_join(tahoe_source_ti, by = "drug_key") %>%
    left_join(tahoe_target_ti, by = c("target_key", "drug_key")) %>%
    mutate(delta_TI = target_TI - source_TI)

  tahoe_moa_ti_summary_df <- tahoe_moa_ti_df %>%
    filter(
      is.finite(delta_NES),
      is.finite(source_TI),
      is.finite(target_TI),
      is.finite(delta_TI)
    ) %>%
    group_by(regime, moa_broad_group, `moa-fine`, moa_fine_label) %>%
    summarise(
      median_delta_NES = median(delta_NES, na.rm = TRUE),
      median_source_TI = median(source_TI, na.rm = TRUE),
      median_target_TI = median(target_TI, na.rm = TRUE),
      median_delta_TI = median(delta_TI, na.rm = TRUE),
      n_drugs = n_distinct(gene),
      .groups = "drop"
    )

  saveRDS(
    tahoe_moa_ti_summary_df,
    file.path(RESULTS_PATH, "tahoe/tahoe_moa_ti_quadrant_summary_df.rds")
  )
} else {
  tahoe_moa_ti_summary_df <- readRDS(file.path(RESULTS_PATH, "tahoe/tahoe_moa_ti_quadrant_summary_df.rds"))
}

plot_moa_ti_quadrant <- function(df, x_col, x_lab, title) {
  moa_legend_order <- df %>%
    filter(regime == "90th") %>%
    arrange(desc(median_delta_NES)) %>%
    pull(`moa-fine`)
  moa_legend_order <- c(moa_legend_order, setdiff(unique(df$`moa-fine`), moa_legend_order))
  
  plot_df <- df %>%
    mutate(`moa-fine` = factor(`moa-fine`, levels = moa_legend_order))
  
  size_breaks <- function(lims) {
    candidates <- c(1, 2, 3, 5, 7, 10, 15, 20, 30, 50)
    candidates[candidates >= ceiling(lims[1]) & candidates <= floor(lims[2])]
  }
  fit_stats <- df %>%
    filter(is.finite(.data[[x_col]]), is.finite(median_delta_NES)) %>%
    group_by(regime) %>%
    group_modify(function(dat, key) {
      dat <- dat %>% mutate(.x_value = .data[[x_col]])
      if (nrow(dat) < 3 || n_distinct(dat$.x_value) < 2 || n_distinct(dat$median_delta_NES) < 2) {
        return(tibble(label = "R2 = NA\np = NA"))
      }
      fit <- lm(median_delta_NES ~ .x_value, data = dat)
      fit_summary <- summary(fit)
      tibble(
        label = sprintf(
          "R2 = %.2f\np = %.2g",
          fit_summary$r.squared,
          coef(fit_summary)[".x_value", "Pr(>|t|)"]
        )
      )
    }) %>%
    ungroup()
  
  ggplot(plot_df, aes(x = .data[[x_col]], y = median_delta_NES)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey55", linewidth = 0.45) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey55", linewidth = 0.45) +
    geom_hline(yintercept = 0, color = "grey80", linewidth = 0.35) +
    geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.55) +
    geom_point(aes(size = n_drugs, fill = `moa-fine`), shape = 21, color = "grey20", alpha = 0.82) +
    geom_text(
      data = fit_stats,
      aes(x = Inf, y = Inf, label = label),
      inherit.aes = FALSE,
      hjust = 1.08,
      vjust = 1.2,
      size = 4.2
    ) +
    facet_grid(. ~ regime) +
    scale_size(range = c(2, 8), breaks = size_breaks) +
    labs(
      title = title,
      x = x_lab,
      y = "Median delta NES",
      fill = "MOA",
      size = "N drugs"
    ) +
    theme_bw(base_size = 16) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "right",
      strip.text = element_text(size = 16)
    )
}

source_ti_quadrant <- plot_moa_ti_quadrant(
  tahoe_moa_ti_summary_df,
  "median_source_TI",
  "Median source TI",
  "MOA-level source TI vs delta NES"
)

target_ti_quadrant <- plot_moa_ti_quadrant(
  tahoe_moa_ti_summary_df,
  "median_target_TI",
  "Median target TI",
  "MOA-level target TI vs delta NES"
)

delta_ti_quadrant <- plot_moa_ti_quadrant(
  tahoe_moa_ti_summary_df,
  "median_delta_TI",
  "Median delta TI",
  "MOA-level delta TI vs delta NES"
)
```

```{r, fig.width=14, fig.height=6}
source_ti_quadrant
```

```{r}
ggsave(
  file.path(PROJECT_PATH, "results/plots/tahoe_moa_source_TI_delta_NES_quadrant.png"),
  source_ti_quadrant,
  width = 14,
  height = 6,
  dpi = 400
)
```

```{r, fig.width=14, fig.height=6}
target_ti_quadrant
```

```{r}
ggsave(
  file.path(PROJECT_PATH, "results/plots/tahoe_moa_target_TI_delta_NES_quadrant.png"),
  target_ti_quadrant,
  width = 14,
  height = 6,
  dpi = 400
)
```

```{r, fig.width=14, fig.height=6}
delta_ti_quadrant
```

```{r}
ggsave(
  file.path(PROJECT_PATH, "results/plots/tahoe_moa_delta_TI_delta_NES_quadrant.png"),
  delta_ti_quadrant,
  width = 14,
  height = 6,
  dpi = 400
)
```

# Paper Figures

```{r paper-fig-setup, message=FALSE, warning=FALSE}
# Shared constants for paper figures
dataset_levels <- c("DrugMatrix", "Perturb-seq", "SciPlex", "Tahoe")

recode_dataset <- function(x) {
  dplyr::recode(x,
    drugmatrix  = "DrugMatrix",
    perturbseq  = "Perturb-seq",
    sciplex     = "SciPlex",
    tahoe       = "Tahoe"
  )
}

# Method label mapping: CSV name -> display label (for Figure 3 heatmap)
method_labels_hm <- c(
  "mean"               = "Mean",
  "project cor (eigen)"= "projCor-Eigen",
  "project cor (gsva)" = "projCor-GSVA",
  "network prop"       = "NetProp",
  "orthos"             = "Orthos",
  "scGPT"              = "scGPT",
  "stack"              = "Stack"
)

# Display order for heatmap y-axis (Mean at top, Stack at bottom)
method_order_csv  <- c("mean", "project cor (eigen)", "project cor (gsva)",
                        "network prop", "scGPT", "stack")
hm_method_levels  <- rev(unname(method_labels_hm[method_order_csv]))
```

## Figure 2: Control regime is a null baseline

```{r paper-fig2, message=FALSE, warning=FALSE, fig.width=14, fig.height=9}
# Load ctrl swarm data for all 4 datasets (both metrics combined).
# Each dataset has a different method subset, so we build per-dataset plots and
# stitch them with cowplot — this avoids empty x-axis slots from missing methods.
ctrl_swarm_all <- dplyr::bind_rows(
  readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_paired_swarm_plot_df.rds")) %>%
    dplyr::mutate(dataset = "DrugMatrix"),
  readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_paired_swarm_plot_df.rds")) %>%
    dplyr::mutate(dataset = "Perturb-seq"),
  readRDS(file.path(RESULTS_PATH, "sciplex/ctrl_paired_swarm_plot_df.rds")) %>%
    dplyr::mutate(dataset = "SciPlex"),
  readRDS(file.path(RESULTS_PATH, "tahoe/ctrl_paired_swarm_plot_df.rds")) %>%
    dplyr::mutate(dataset = "Tahoe")
) %>%
  dplyr::mutate(
    dataset = factor(dataset, levels = dataset_levels),
    method  = dplyr::recode(method,
      "NetworkProp"        = "NetProp",
      "ProjectCor (Eigen)" = "projCor-Eigen",
      "ProjectCor (GSVA)"  = "projCor-GSVA"
    )
  )

# Fixed Pastel1 colors keyed by method name for consistency across panels
method_pal_fig2 <- setNames(
  RColorBrewer::brewer.pal(6, "Pastel1"),
  c("Mean", "NetProp", "projCor-Eigen", "projCor-GSVA", "Orthos", "Stack")
)

# Methods available per dataset (in desired x-axis order)
ds_methods <- list(
  "DrugMatrix"  = c("Mean", "NetProp", "projCor-Eigen", "projCor-GSVA", "Orthos"),
  "Perturb-seq" = c("Mean", "NetProp", "projCor-Eigen", "projCor-GSVA", "Stack"),
  "SciPlex"     = c("Mean", "NetProp", "projCor-Eigen", "projCor-GSVA", "Orthos", "Stack"),
  "Tahoe"       = c("Mean", "NetProp", "projCor-Eigen", "projCor-GSVA", "Orthos", "Stack")
)

# One panel per dataset, one row per metric
make_ds_panel <- function(ds, metric_name,
                           show_title = TRUE, show_ylab = TRUE) {
  methods <- ds_methods[[ds]]
  dat <- ctrl_swarm_all %>%
    dplyr::filter(dataset == ds, metric == metric_name) %>%
    dplyr::mutate(method = factor(method, levels = methods))

  p <- ggplot(dat, aes(x = method, y = value, fill = method)) +
    geom_violin(width = 0.9, alpha = 0.35, color = NA) +
    ggbeeswarm::geom_quasirandom(width = 0.15, alpha = 0.15, size = 0.05,
                                  groupOnX = TRUE) +
    stat_summary(fun = median, geom = "point", shape = 21, size = 1,
                 fill = "white") +
    geom_hline(yintercept = 0, linetype = "dotted", color = "grey50",
               linewidth = 0.8) +
    scale_fill_manual(values = method_pal_fig2) +
    scale_x_discrete(limits = methods) +
    labs(x = "",
         y = if (show_ylab) "Delta NES" else "") +
    theme_classic(base_size = 9) +
    theme(axis.text.x     = element_text(angle = 35, hjust = 1),
          legend.position = "none")

  if (show_title)
    p <- p + ggtitle(ds) +
      theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 11))
  if (!show_ylab)
    p <- p + theme(axis.text.y  = element_blank(),
                   axis.ticks.y = element_blank())
  p
}

# Row A: NES∆  — dataset titles shown, y-label on left panel only
row_nes <- cowplot::plot_grid(
  make_ds_panel("DrugMatrix",  "NES \u2206", show_title = TRUE,  show_ylab = TRUE),
  make_ds_panel("Perturb-seq", "NES \u2206", show_title = TRUE,  show_ylab = FALSE),
  make_ds_panel("SciPlex",     "NES \u2206", show_title = TRUE,  show_ylab = FALSE),
  make_ds_panel("Tahoe",       "NES \u2206", show_title = TRUE,  show_ylab = FALSE),
  nrow = 1, align = "hv", axis = "tb",
  rel_widths = c(5, 5, 6, 6)   # proportional to method count
)

# Row B: Jaccard∆ — no repeated titles, y-label updated
make_ds_panel_jacc <- function(ds, show_ylab) {
  methods <- ds_methods[[ds]]
  dat <- ctrl_swarm_all %>%
    dplyr::filter(dataset == ds, metric == "Jaccard \u2206") %>%
    dplyr::mutate(method = factor(method, levels = methods))

  p <- ggplot(dat, aes(x = method, y = value, fill = method)) +
    geom_violin(width = 0.9, alpha = 0.35, color = NA) +
    ggbeeswarm::geom_quasirandom(width = 0.15, alpha = 0.15, size = 0.05,
                                  groupOnX = TRUE) +
    stat_summary(fun = median, geom = "point", shape = 21, size = 1,
                 fill = "white") +
    geom_hline(yintercept = 0, linetype = "dotted", color = "grey50",
               linewidth = 0.8) +
    scale_fill_manual(values = method_pal_fig2) +
    scale_x_discrete(limits = methods) +
    labs(x = "", y = if (show_ylab) "Delta Jaccard" else "") +
    theme_classic(base_size = 9) +
    theme(axis.text.x     = element_text(angle = 35, hjust = 1),
          legend.position = "none")

  if (!show_ylab)
    p <- p + theme(axis.text.y  = element_blank(),
                   axis.ticks.y = element_blank())
  p
}

row_jacc <- cowplot::plot_grid(
  make_ds_panel_jacc("DrugMatrix",  show_ylab = TRUE),
  make_ds_panel_jacc("Perturb-seq", show_ylab = FALSE),
  make_ds_panel_jacc("SciPlex",     show_ylab = FALSE),
  make_ds_panel_jacc("Tahoe",       show_ylab = FALSE),
  nrow = 1, align = "hv", axis = "tb",
  rel_widths = c(5, 5, 6, 6)
)

fig2 <- cowplot::plot_grid(row_nes, row_jacc, ncol = 1,
                            labels = c("A", "B"), label_size = 11,
                            rel_heights = c(1.15, 1))
fig2
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/paper_fig2_ctrl_null.pdf"),
       fig2, width = 8, height = 5.14, dpi = 400, device = cairo_pdf)
```

## Figure 3: Information scaling and method landscape

```{r paper-fig3, message=FALSE, warning=FALSE, fig.width=12, fig.height=9}
# Each cell: fill = -log10(Fisher NES/Jacc meta-p); label = rank by median W
# statistic within each dataset column. Rows = regime, columns = metric.
# Shows: (1) which datasets respond to target data, (2) no single method wins.

build_regime_hm <- function(paired_p, paired_stat, regime_label) {
  # p-values → long with neg_log10p
  p_long <- paired_p %>%
    tidyr::pivot_longer(-method, names_to = "col", values_to = "p") %>%
    tidyr::separate(col, into = c("dataset", "metric"), sep = "_(?=[^_]+$)") %>%
    dplyr::filter(!is.na(p)) %>%
    dplyr::mutate(
      neg_log10p = pmin(-log10(pmax(p, 1e-30)), 30),
      dataset    = factor(recode_dataset(dataset), levels = dataset_levels),
      metric     = dplyr::recode(metric, NES = "NES", Jacc = "Jaccard"),
      metric     = factor(metric, levels = c("NES", "Jaccard")),
      method     = dplyr::recode(method, !!!method_labels_hm),
      method     = factor(method, levels = hm_method_levels)
    )

  # W statistics → rank within each dataset × metric
  rank_long <- paired_stat %>%
    tidyr::pivot_longer(-method, names_to = "col", values_to = "W") %>%
    tidyr::separate(col, into = c("dataset", "metric"), sep = "_(?=[^_]+$)") %>%
    dplyr::filter(!is.na(W)) %>%
    dplyr::mutate(
      dataset = factor(recode_dataset(dataset), levels = dataset_levels),
      metric  = dplyr::recode(metric, NES = "NES", Jacc = "Jaccard"),
      metric  = factor(metric, levels = c("NES", "Jaccard")),
      method  = dplyr::recode(method, !!!method_labels_hm),
      method  = factor(method, levels = hm_method_levels)
    ) %>%
    dplyr::group_by(dataset, metric) %>%
    dplyr::mutate(rank = rank(-W, ties.method = "min")) %>%
    dplyr::ungroup() %>%
    dplyr::select(method, dataset, metric, rank)

  p_long %>%
    dplyr::left_join(rank_long, by = c("method", "dataset", "metric")) %>%
    dplyr::mutate(regime = regime_label)
}

fig3_df_raw <- dplyr::bind_rows(
  build_regime_hm(paired_10th, paired_stat_10th,
                  "Resource constrained (1/10)"),
  build_regime_hm(paired_90th, paired_stat_90th,
                  "Maximal information (9/10)")
)

# Fill in all method × dataset × metric × regime combinations so that cells
# with no eval data (e.g. scGPT for DrugMatrix/SciPlex/Tahoe, Stack for
# DrugMatrix) are explicitly represented with NA fill and "N/A" label.
all_combos <- tidyr::expand_grid(
  method  = factor(hm_method_levels, levels = hm_method_levels),
  dataset = factor(dataset_levels,   levels = dataset_levels),
  metric  = factor(c("NES","Jaccard"), levels = c("NES","Jaccard")),
  regime  = c("Resource constrained (1/10)", "Maximal information (9/10)")
)

fig3_df <- all_combos %>%
  dplyr::left_join(fig3_df_raw,
                   by = c("method","dataset","metric","regime")) %>%
  dplyr::mutate(
    regime     = factor(regime,
                        levels = c("Resource constrained (1/10)",
                                   "Maximal information (9/10)")),
    rank_label = dplyr::case_when(
      !is.na(rank)        ~ as.character(rank),
      !is.na(neg_log10p)  ~ "",        # p present but rank absent (shouldn't occur)
      TRUE                ~ "N/A"      # no eval data at all
    ),
    label_color = ifelse(!is.na(rank) & rank == 1, "red1", "gray15")
  )

fig3 <- ggplot(fig3_df, aes(x = dataset, y = method, fill = neg_log10p)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = rank_label, color = label_color), size = 4, fontface = "bold") +
  scale_color_identity() +
  facet_grid(regime ~ metric) +
  scale_fill_distiller(palette = "GnBu", direction = 1,
                       name = "-log₁₀(p)", limits = c(0, 30),
                       na.value = "grey90") +
  labs(x = NULL, y = NULL) +
  theme_classic(base_size = 9) +
  theme(
    axis.text.x     = element_text(angle = 35, hjust = 1),
    strip.text      = element_text(face = "bold", size = 9),
    panel.grid      = element_blank(),
    legend.position = "right"
  )
fig3
```

```{r paper-fig3c, message=FALSE, warning=FALSE}
# Panel B: predicted vs source scatter for best method per dataset (90th regime).
# Points are colored by target context only (legend title = "Target").
# Target label derivation:
#   DrugMatrix  — target column is already a single context name ("liver"/"kidney")
#   Perturb-seq — target column is placeholder; parse target from source ("k562_rpe1" → "RPE1")
#   SciPlex     — same parsing as Perturb-seq
#   Tahoe       — target column already nicely formatted (cell line names)

qual_pal_paired <- RColorBrewer::brewer.pal(9, "Set1")

# Extract target label from compound source string ("k562_rpe1" → "RPE1")
fmt_target_from_source <- function(source) {
  toupper(sapply(strsplit(source, "_"), `[[`, 2))
}

# Per-dataset specs for NES version (best by W statistic)
best_90th_NES <- list(
  list(title    = "DrugMatrix (Best Method: Mean)",
       file     = "drugmatrix/ctrl_90th_mean_eval_table.rds",
       tgt_fn   = function(df) tools::toTitleCase(df$target),
       p_method = "mean",               p_col = "drugmatrix_NES"),
  list(title    = "Perturb-seq (Best Method: NetProp)",
       file     = "perturb-seq/ctrl_90th_networkprop_eval_table.rds",
       tgt_fn   = function(df) fmt_target_from_source(df$source),
       p_method = "network prop",       p_col = "perturbseq_NES"),
  list(title    = "SciPlex (Best Method: projCor-GSVA)",
       file     = "sciplex/90th_projectcor_gsva_eval_table.rds",
       tgt_fn   = function(df) fmt_target_from_source(df$source),
       p_method = "project cor (gsva)", p_col = "sciplex_NES"),
  list(title    = "Tahoe (Best Method: NetProp)",
       file     = "tahoe/network_prop_90th_eval_table.rds",
       tgt_fn   = function(df) df$target,
       p_method = "network prop",       p_col = "tahoe_NES")
)

# Per-dataset specs for Jaccard version (projCor-GSVA best for PS and SC)
best_90th_Jacc <- list(
  list(title  = "DrugMatrix (Best Method: Mean)",
       file   = "drugmatrix/ctrl_90th_mean_eval_table.rds",
       tgt_fn = function(df) tools::toTitleCase(df$target)),
  list(title  = "Perturb-seq (Best Method: projCor-GSVA)",
       file   = "perturb-seq/90th_projectcor_gsva_eval_table.rds",
       tgt_fn = function(df) fmt_target_from_source(df$source)),
  list(title  = "SciPlex (Best Method: projCor-GSVA)",
       file   = "sciplex/90th_projectcor_gsva_eval_table.rds",
       tgt_fn = function(df) fmt_target_from_source(df$source)),
  list(title  = "Tahoe (Best Method: NetProp)",
       file   = "tahoe/network_prop_90th_eval_table.rds",
       tgt_fn = function(df) df$target)
)

make_scatter <- function(item, x_col, y_col, xlab, ylab) {
  df <- readRDS(file.path(RESULTS_PATH, item$file)) %>%
    dplyr::filter(!is.na(.data[[y_col]]), !is.na(.data[[x_col]]),
                  is.finite(.data[[y_col]]), is.finite(.data[[x_col]])) %>%
    dplyr::mutate(target_ctx = item$tgt_fn(.))
  lims   <- range(c(df[[x_col]], df[[y_col]]))
  pad    <- diff(lims) * 0.03
  lims_p <- c(lims[1] - pad, lims[2] + pad)

  subtitle <- if (!is.null(item$p_method)) {
    p_val <- paired_90th %>% dplyr::filter(method == item$p_method) %>% dplyr::pull(item$p_col)
    if (is.na(p_val) || p_val >= 0.05) {
      "Wilcox meta p-value: NS"
    } else {
      paste0("Wilcox meta p-value: ", formatC(p_val, format = "e", digits = 2))
    }
  } else NULL

  ggplot(df, aes(x = .data[[x_col]], y = .data[[y_col]], color = target_ctx)) +
    geom_point(alpha = 0.5, size = 0.8) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted",
                color = "grey40", linewidth = 0.7) +
    scale_color_manual(values = qual_pal_paired, name = "Target") +
    scale_x_continuous(limits = lims_p) +
    scale_y_continuous(limits = lims_p) +
    labs(title = item$title, subtitle = subtitle, x = xlab, y = ylab) +
    theme_classic(base_size = 9) +
    theme(plot.title      = element_text(face = "bold", hjust = 0.5, size = 8),
          plot.subtitle   = element_text(hjust = 0.5, size = 7, color = "gray30"),
          legend.position  = "right",
          legend.title     = element_text(size = 10),
          legend.text      = element_text(size = 7),
          legend.key.size  = unit(0.5, "lines"),
          aspect.ratio     = 1)
}

build_panel_b <- function(best_list, x_col, y_col, xlab, ylab) {
  plots <- lapply(best_list, make_scatter,
                  x_col = x_col, y_col = y_col, xlab = xlab, ylab = ylab)
  cowplot::plot_grid(plots[[1]], plots[[2]], plots[[3]], plots[[4]], nrow = 2)
}

panel_b_NES  <- build_panel_b(best_90th_NES,  "NES_FALSE",  "NES_TRUE",
                               "NES (source)", "NES (predicted)")
panel_b_Jacc <- build_panel_b(best_90th_Jacc, "jacc_FALSE", "jacc_TRUE",
                               "Jaccard (source)", "Jaccard (predicted)")

fig3_NES <- cowplot::plot_grid(fig3, panel_b_NES,
                                ncol = 1, labels = c("A", "B"), label_size = 11,
                                rel_heights = c(1, 1))
fig3_NES
```

```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/paper_fig3_NES.pdf"),
       fig3_NES,    width = 8, height = 11, dpi = 400, device = cairo_pdf)
ggsave(file.path(PROJECT_PATH, "results/plots/paper_fig3_Jacc.pdf"),
       panel_b_Jacc, width = 7, height = 7,  dpi = 400, device = cairo_pdf)
```


# TI Distributions

```{r ti-distributions, message=FALSE, warning=FALSE}
ti_path <- file.path(PROJECT_PATH, "results/plots/trade_metric_correlations")

ti_df <- bind_rows(
  readRDS(file.path(ti_path, "drugmatrix/context_trade_metrics.rds")) %>%
    transmute(dataset = "DrugMatrix", TI),
  readRDS(file.path(ti_path, "perturb_seq/context_trade_metrics.rds")) %>%
    transmute(dataset = "Perturb-seq", TI),
  readRDS(file.path(ti_path, "sciplex/context_trade_metrics.rds")) %>%
    transmute(dataset = "SciPlex", TI),
  readRDS(file.path(ti_path, "tahoe/context_trade_metrics.rds")) %>%
    transmute(dataset = "Tahoe", TI)
) %>%
  filter(is.finite(TI)) %>%
  mutate(dataset = factor(dataset, levels = c("DrugMatrix", "Perturb-seq", "SciPlex", "Tahoe")))
```

```{r}
# ECDF — shaded area + step line per dataset, log1p x-axis.
# stat_ecdf called twice: once as "area" for fill, once as "step" for the line.
# The jump at x=0 shows the zero-inflation fraction for each dataset.
ggplot(ti_df, aes(x = log1p(TI), color = dataset)) +
  stat_ecdf(geom = "line", alpha = 0.35, linewidth = 0) +
  stat_ecdf(linewidth = 0.9) +
  scale_x_continuous(
    name   = "TI",
    breaks = log1p(c(0, 0.1, 0.25, 0.5, 1, 2, 3)),
    labels = c("0", "0.1", "0.25", "0.5", "1", "2", "3")
  ) +
  scale_y_continuous(name = "Cumulative fraction", labels = scales::percent) +
  scale_color_brewer(palette = "Paired", name = "Dataset") +
  theme_classic(base_size = 15) +
  theme(legend.position = "right")
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/ti_distributions_ecdf.png"),
       width = 8, height = 5, dpi = 400)
```

```{r}
# Complementary CDF — shaded ribbon below curve, log-log axes.
# Reads as "fraction of perturbations with TI > x".
# Log-log reveals tail heaviness; a straight line indicates a power-law tail.
# ymin = 1e-4 anchors the ribbon to the bottom of the log y-axis.
ti_ccdf_df <- ti_df %>%
  group_by(dataset) %>%
  arrange(TI, .by_group = TRUE) %>%
  mutate(ccdf = 1 - (seq_len(n()) / n())) %>%
  ungroup()

p_ccdf <- ggplot(ti_ccdf_df %>% filter(TI > 0),
       aes(x = TI, color = dataset)) +
  geom_line(aes(y = ccdf), linewidth = 0.9) +
  # scale_x_log10(name = "TI") +
  scale_y_log10(name = "Fraction with TI > x", labels = scales::percent) +
  scale_color_brewer(palette = "Paired", name = "Dataset") +
  theme_classic(base_size = 9) +
  theme(legend.position          = c(0.98, 0.98),
        legend.justification     = c("right", "top"),
        legend.background        = element_rect(fill = alpha("white", 0.7), colour = NA),
        legend.key.size          = unit(0.5, "lines"))
p_ccdf
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/ti_distributions_ccdf.png"),
       width = 8, height = 5, dpi = 400)
```

```{r}
# Proportional bar chart by TI category.
# Bins: zero, low (0, 0.1], medium (0.1, 0.5], high (> 0.5)
ti_cat_df <- ti_df %>%
  mutate(ti_cat = case_when(
    TI <  0.1 ~ "Low [0–0.1)",
    TI <  0.5 ~ "Medium [0.1–0.5)",
    TRUE      ~ "High (≥0.5)"
  )) %>%
  mutate(ti_cat = factor(ti_cat,
                         levels = c("Low [0–0.1)", "Medium [0.1–0.5)",
                                    "High (≥0.5)")))

ti_cat_summary <- ti_cat_df %>%
  count(dataset, ti_cat) %>%
  group_by(dataset) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

p_bar <- ggplot(ti_cat_summary, aes(x = dataset, y = prop, fill = ti_cat)) +
  geom_col(width = 0.7) +
  scale_y_continuous(name = "Fraction of perturbations", labels = scales::percent) +
  scale_fill_brewer(palette = "Paired", name = "TI category") +
  labs(x = "") +
  theme_classic(base_size = 9) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        legend.position = "right")
p_bar
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/ti_distributions_categories.png"),
       width = 7, height = 5, dpi = 400)
```

## Figure 5: TI × Delta NES (90th regime, best method per dataset)

```{r ti-delta-nes, message=FALSE, warning=FALSE}
# Best method per dataset by median NES Wilcoxon W statistic in the 90th regime:
#   DrugMatrix  → mean             (W = 643.5)
#   Perturb-seq → network prop     (W = 554,545)
#   SciPlex     → project cor gsva (W = 218)
#   Tahoe       → network prop     (W = 177,611)

ti_delta_eval <- bind_rows(
  readRDS(file.path(RESULTS_PATH, "drugmatrix/ctrl_90th_mean_eval_table.rds")) %>%
    transmute(dataset = "DrugMatrix", source, gene,
              delta_NES = NES_TRUE - NES_FALSE),
  readRDS(file.path(RESULTS_PATH, "perturb-seq/ctrl_90th_networkprop_eval_table.rds")) %>%
    transmute(dataset = "Perturb-seq", source, gene,
              delta_NES = NES_TRUE - NES_FALSE),
  readRDS(file.path(RESULTS_PATH, "sciplex/90th_projectcor_gsva_eval_table.rds")) %>%
    transmute(dataset = "SciPlex", source, gene,
              delta_NES = NES_TRUE - NES_FALSE),
  readRDS(file.path(RESULTS_PATH, "tahoe/network_prop_90th_eval_table.rds")) %>%
    transmute(dataset = "Tahoe", source, gene,
              delta_NES = NES_TRUE - NES_FALSE)
)

# Source-context TI for all datasets
ti_all <- bind_rows(
  readRDS(file.path(ti_path, "drugmatrix/context_trade_metrics.rds")) %>%
    transmute(dataset = "DrugMatrix", context_key, drug, TI),
  readRDS(file.path(ti_path, "perturb_seq/context_trade_metrics.rds")) %>%
    transmute(dataset = "Perturb-seq", context_key, drug, TI),
  readRDS(file.path(ti_path, "sciplex/context_trade_metrics.rds")) %>%
    transmute(dataset = "SciPlex", context_key, drug, TI),
  readRDS(file.path(ti_path, "tahoe/context_trade_metrics.rds")) %>%
    transmute(dataset = "Tahoe", context_key, drug, TI)
)

# Normalize eval `source` to match TI `context_key`:
#   Tahoe       "nci_h23"   → "ncih23"  (strip underscores)
#   SciPlex     "a549_k562" → "a549"    (take component before first _)
#   Perturb-seq "k562_rpe1" → "k562"   (take component before first _; network_prop uses source_target format)
#   DrugMatrix  unchanged   (e.g. "kidney")
ti_delta_joined <- ti_delta_eval %>%
  mutate(
    source_key = case_when(
      dataset == "Tahoe"       ~ gsub("_", "", source),
      dataset == "SciPlex"     ~ sub("_.*", "", source),
      dataset == "Perturb-seq" ~ sub("_.*", "", source),
      TRUE                     ~ source
    )
  ) %>%
  left_join(ti_all,
            by = c("dataset", "source_key" = "context_key", "gene" = "drug")) %>%
  filter(!is.na(TI), !is.na(delta_NES)) %>%
  mutate(
    ti_cat = case_when(
      TI <  0.1 ~ "Low [0–0.1)",
      TI <  0.5 ~ "Medium [0.1–0.5)",
      TRUE      ~ "High (≥0.5)"
    ),
    ti_cat  = factor(ti_cat, levels = c("Low [0–0.1)", "Medium [0.1–0.5)",
                                         "High (≥0.5)")),
    dataset = factor(dataset,
                     levels = c("DrugMatrix", "Perturb-seq", "SciPlex", "Tahoe"))
  )
```

```{r}
# Pre-compute two-sided Wilcox p-values per dataset for each TI comparison pair.
# Only retain p < 0.05; use ggsignif manual mode to avoid the ggpubr/ggsignif
# bug where hide.ns = TRUE suppresses the label but still draws the bracket line.
ti_comp_pairs <- tibble::tribble(
  ~higher,              ~lower,
  "Medium [0.1–0.5)", "Low [0–0.1)",
  "High (≥0.5)",      "Medium [0.1–0.5)",
  "High (≥0.5)",      "Low [0–0.1)"
)

ti_signif_df <- purrr::map_dfr(
  unique(ti_delta_joined$dataset),
  function(ds) {
    dat <- filter(ti_delta_joined, dataset == ds)
    purrr::pmap_dfr(ti_comp_pairs, function(higher, lower) {
      x <- dat %>% filter(ti_cat == higher) %>% pull(delta_NES)
      y <- dat %>% filter(ti_cat == lower)  %>% pull(delta_NES)
      if (length(x) < 3 || length(y) < 3) return(NULL)
      p <- wilcox.test(x, y)$p.value
      tibble(dataset = ds,
             xmin = lower, xmax = higher,   # left→right bracket positions on x-axis
             p_value = p)
    })
  }
) %>%
  filter(p_value < 0.05) %>%
  mutate(
    annotations = case_when(
      p_value < 0.0001 ~ "****",
      p_value < 0.001  ~ "***",
      p_value < 0.01   ~ "**",
      TRUE             ~ "*"
    ),
    dataset = factor(dataset, levels = c("DrugMatrix", "Perturb-seq", "SciPlex", "Tahoe"))
  )

# Compute per-dataset y-positions for stacked brackets (data-scale, free_y safe)
y_stats <- ti_delta_joined %>%
  group_by(dataset) %>%
  summarize(y_max   = max(delta_NES, na.rm = TRUE),
            y_range = diff(range(delta_NES, na.rm = TRUE)),
            .groups = "drop")

ti_signif_df <- ti_signif_df %>%
  left_join(y_stats, by = "dataset") %>%
  group_by(dataset) %>%
  mutate(y_position = y_max + row_number() * y_range * 0.13) %>%
  ungroup()

ti_cat_levels_ordered <- c("Low [0–0.1)", "Medium [0.1–0.5)", "High (≥0.5)")
signif_annot <- ti_signif_df %>%
  mutate(
    x1   = match(xmin, ti_cat_levels_ordered),
    x2   = match(xmax, ti_cat_levels_ordered),
    xmid = (x1 + x2) / 2,
    tip  = y_range * 0.025
  )

p_delta <- ggplot(ti_delta_joined, aes(x = ti_cat, y = delta_NES, fill = ti_cat)) +
  geom_boxplot(outlier.size = 0.4, outlier.alpha = 0.3, linewidth = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.7) +
  geom_segment(data = signif_annot,
    aes(x = x1, xend = x2, y = y_position, yend = y_position),
    inherit.aes = FALSE, linewidth = 0.4) +
  geom_segment(data = signif_annot,
    aes(x = x1, xend = x1, y = y_position - tip, yend = y_position),
    inherit.aes = FALSE, linewidth = 0.4) +
  geom_segment(data = signif_annot,
    aes(x = x2, xend = x2, y = y_position - tip, yend = y_position),
    inherit.aes = FALSE, linewidth = 0.4) +
  geom_text(data = signif_annot,
    aes(x = xmid, y = y_position, label = annotations),
    inherit.aes = FALSE, size = 3.5, vjust = 0) +
  facet_wrap(~dataset, scales = "free_y", ncol = 2) +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
  labs(x = "TI category", y = "Delta NES (predicted − source)") +
  theme_classic(base_size = 9) +
  theme(
    axis.text.x   = element_text(angle = 35, hjust = 1),
    legend.position = "none",
    strip.text    = element_text(face = "bold"),
    panel.spacing = unit(1.5, "lines")
  )
p_delta
```
```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/ti_delta_nes_by_category.png"),
       width = 9, height = 7, dpi = 400)
```

```{r ti-combined-panel}
# Row 1: CCDF (left) + stacked bar chart (right)
# Row 2: Delta NES × TI boxplot (full width)
top_row   <- cowplot::plot_grid(p_ccdf, p_bar, nrow = 1, labels = c("A", "B"),
                                label_size = 11, rel_widths = c(1, 0.8))
panel_fig <- cowplot::plot_grid(top_row, p_delta, ncol = 1, labels = c("", "C"),
                                label_size = 11, rel_heights = c(1, 1.2))
panel_fig
```

```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/paper_fig5_combined_panel.pdf"),
       panel_fig, width = 8, height = 8, dpi = 400, device = "pdf")
```

# Fig 5b: pi_DEG × Delta NES

```{r fig5b-pideg, message=FALSE, warning=FALSE}
# Global 33rd/67th percentile cuts for pi_DEG across all four datasets,
# rounded to nearest 1000: Low <2k, Medium 2k-4k, High >=4k genes.
PI_LOW  <- 2000
PI_HIGH <- 4000

pideg_cat_levels <- c("Low (<2k)", "Medium (2k–4k)", "High (≥4k)")

pideg_all <- bind_rows(
  readRDS(file.path(ti_path, "drugmatrix/context_trade_metrics.rds"))  %>% transmute(dataset = "DrugMatrix",  TI, pi_DEG),
  readRDS(file.path(ti_path, "perturb_seq/context_trade_metrics.rds")) %>% transmute(dataset = "Perturb-seq", TI, pi_DEG),
  readRDS(file.path(ti_path, "sciplex/context_trade_metrics.rds"))     %>% transmute(dataset = "SciPlex",     TI, pi_DEG),
  readRDS(file.path(ti_path, "tahoe/context_trade_metrics.rds"))       %>% transmute(dataset = "Tahoe",       TI, pi_DEG)
) %>%
  filter(is.finite(pi_DEG)) %>%
  mutate(
    dataset   = factor(dataset, levels = c("DrugMatrix", "Perturb-seq", "SciPlex", "Tahoe")),
    pideg_cat = case_when(
      pi_DEG <  PI_LOW  ~ pideg_cat_levels[1],
      pi_DEG <  PI_HIGH ~ pideg_cat_levels[2],
      TRUE              ~ pideg_cat_levels[3]
    ),
    pideg_cat = factor(pideg_cat, levels = pideg_cat_levels)
  )

# Panel A: CCDF of pi_DEG (log x-axis)
pideg_ccdf_df <- pideg_all %>%
  group_by(dataset) %>%
  arrange(pi_DEG, .by_group = TRUE) %>%
  mutate(ccdf = 1 - (seq_len(n()) / n())) %>%
  ungroup()

p_pideg_ccdf <- ggplot(pideg_ccdf_df, aes(x = pi_DEG, color = dataset)) +
  geom_line(aes(y = ccdf), linewidth = 0.9) +
  scale_x_log10(name = expression(pi[DEG] ~ "(log scale)"),
                labels = function(x) paste0("1e", as.integer(log10(x)))) +
  scale_y_log10(name = expression("Fraction with " ~ pi[DEG] ~ "> x"), labels = scales::percent) +
  scale_color_brewer(palette = "Paired", name = "Dataset") +
  theme_classic(base_size = 9) +
  theme(legend.position = "right",
        legend.key.size = unit(0.5, "lines"),
        axis.text.x     = element_text(angle = 45, hjust = 1))

# Panel B: stacked bar of category fractions
pideg_cat_summary <- pideg_all %>%
  count(dataset, pideg_cat) %>%
  group_by(dataset) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

p_pideg_bar <- ggplot(pideg_cat_summary, aes(x = dataset, y = prop, fill = pideg_cat)) +
  geom_col(width = 0.7) +
  scale_y_continuous(name = "Fraction of perturbations", labels = scales::percent) +
  scale_fill_brewer(palette = "Paired", name = expression(pi[DEG] ~ "category")) +
  labs(x = "") +
  theme_classic(base_size = 9) +
  theme(axis.text.x    = element_text(angle = 35, hjust = 1),
        legend.position = "right")

# Join pi_DEG to delta NES (same source-key logic as TI figure)
pideg_join_df <- bind_rows(
  readRDS(file.path(ti_path, "drugmatrix/context_trade_metrics.rds"))  %>% transmute(dataset = "DrugMatrix",  context_key, drug, pi_DEG),
  readRDS(file.path(ti_path, "perturb_seq/context_trade_metrics.rds")) %>% transmute(dataset = "Perturb-seq", context_key, drug, pi_DEG),
  readRDS(file.path(ti_path, "sciplex/context_trade_metrics.rds"))     %>% transmute(dataset = "SciPlex",     context_key, drug, pi_DEG),
  readRDS(file.path(ti_path, "tahoe/context_trade_metrics.rds"))       %>% transmute(dataset = "Tahoe",       context_key, drug, pi_DEG)
)

pideg_delta_joined <- ti_delta_eval %>%
  mutate(source_key = case_when(
    dataset == "Tahoe"                       ~ gsub("_", "", source),
    dataset %in% c("SciPlex", "Perturb-seq") ~ sub("_.*", "", source),
    TRUE                                     ~ source
  )) %>%
  left_join(pideg_join_df, by = c("dataset", "source_key" = "context_key", "gene" = "drug")) %>%
  filter(!is.na(pi_DEG), !is.na(delta_NES), is.finite(pi_DEG)) %>%
  mutate(
    pideg_cat = case_when(
      pi_DEG <  PI_LOW  ~ pideg_cat_levels[1],
      pi_DEG <  PI_HIGH ~ pideg_cat_levels[2],
      TRUE              ~ pideg_cat_levels[3]
    ),
    pideg_cat = factor(pideg_cat, levels = pideg_cat_levels),
    dataset   = factor(dataset, levels = c("DrugMatrix", "Perturb-seq", "SciPlex", "Tahoe"))
  )

# Significance brackets (Wilcoxon, p < 0.05)
pideg_comp_pairs <- tibble::tribble(
  ~higher,             ~lower,
  pideg_cat_levels[2], pideg_cat_levels[1],
  pideg_cat_levels[3], pideg_cat_levels[2],
  pideg_cat_levels[3], pideg_cat_levels[1]
)

pideg_signif_df <- purrr::map_dfr(unique(pideg_delta_joined$dataset), function(ds) {
  dat <- filter(pideg_delta_joined, dataset == ds)
  purrr::pmap_dfr(pideg_comp_pairs, function(higher, lower) {
    x <- dat %>% filter(pideg_cat == higher) %>% pull(delta_NES)
    y <- dat %>% filter(pideg_cat == lower)  %>% pull(delta_NES)
    if (length(x) < 3 || length(y) < 3) return(NULL)
    tibble(dataset = ds, xmin = lower, xmax = higher, p_value = wilcox.test(x, y)$p.value)
  })
}) %>%
  filter(p_value < 0.05) %>%
  mutate(
    annotations = case_when(
      p_value < 0.0001 ~ "****",
      p_value < 0.001  ~ "***",
      p_value < 0.01   ~ "**",
      TRUE             ~ "*"
    ),
    dataset = factor(dataset, levels = c("DrugMatrix", "Perturb-seq", "SciPlex", "Tahoe"))
  )

pideg_y_stats <- pideg_delta_joined %>%
  group_by(dataset) %>%
  summarize(y_max   = max(delta_NES, na.rm = TRUE),
            y_range = diff(range(delta_NES, na.rm = TRUE)),
            .groups = "drop")

pideg_signif_df <- pideg_signif_df %>%
  left_join(pideg_y_stats, by = "dataset") %>%
  group_by(dataset) %>%
  mutate(y_position = y_max + row_number() * y_range * 0.13) %>%
  ungroup()

pideg_signif_annot <- pideg_signif_df %>%
  mutate(
    x1   = match(xmin, pideg_cat_levels),
    x2   = match(xmax, pideg_cat_levels),
    xmid = (x1 + x2) / 2,
    tip  = y_range * 0.025
  )

# Panel C: Delta NES boxplot by pi_DEG category — one row, n labels on x-axis
n_per_cat_c <- pideg_delta_joined %>%
  count(dataset, pideg_cat)

pideg_delta_joined <- pideg_delta_joined %>%
  left_join(n_per_cat_c, by = c("dataset", "pideg_cat")) %>%
  mutate(
    pideg_cat_n = paste0(as.character(pideg_cat), "\n(n=", n, ")"),
    pideg_cat_n = reorder(pideg_cat_n, as.integer(pideg_cat))
  )

p_pideg_delta <- ggplot(pideg_delta_joined, aes(x = pideg_cat_n, y = delta_NES, fill = pideg_cat)) +
  geom_boxplot(outlier.size = 0.4, outlier.alpha = 0.3, linewidth = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.7) +
  geom_segment(data = pideg_signif_annot,
    aes(x = x1, xend = x2, y = y_position, yend = y_position),
    inherit.aes = FALSE, linewidth = 0.4) +
  geom_segment(data = pideg_signif_annot,
    aes(x = x1, xend = x1, y = y_position - tip, yend = y_position),
    inherit.aes = FALSE, linewidth = 0.4) +
  geom_segment(data = pideg_signif_annot,
    aes(x = x2, xend = x2, y = y_position - tip, yend = y_position),
    inherit.aes = FALSE, linewidth = 0.4) +
  geom_text(data = pideg_signif_annot,
    aes(x = xmid, y = y_position, label = annotations),
    inherit.aes = FALSE, size = 3.5, vjust = 0) +
  facet_wrap(~dataset, scales = "free", nrow = 1) +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
  labs(x = expression(pi[DEG] ~ "category"), y = "Delta NES (predicted − source)") +
  theme_classic(base_size = 9) +
  theme(axis.text.x    = element_text(angle = 45, hjust = 1),
        legend.position = "none",
        strip.text      = element_text(face = "bold"),
        panel.spacing   = unit(1, "lines"))

top_row_pideg   <- cowplot::plot_grid(p_pideg_ccdf, p_pideg_bar, nrow = 1,
                                      labels = c("A", "B"), label_size = 11,
                                      rel_widths = c(1, 0.8))
panel_fig_pideg <- cowplot::plot_grid(top_row_pideg, p_pideg_delta, ncol = 1,
                                      labels = c("", "C"), label_size = 11,
                                      rel_heights = c(1, 0.9))
panel_fig_pideg
```

```{r}
ggsave(file.path(PROJECT_PATH, "results/plots/paper_fig5_pideg.pdf"),
       panel_fig_pideg, width = 8, height = 6.5, dpi = 400, device = cairo_pdf)
```
